### 2017-08-15 docs/latex/draftB.tex
Documentation now in LaTeX and further bug fixes. Updated README.

\documentclass[letterpaper]{ltxdoc}

\usepackage[hmargin={3 cm,3cm},top=1.5cm, marginpar=3cm]{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\hypersetup{colorlinks, linkcolor=blue, urlcolor=blue}
\usepackage{multicol}
\usepackage{lipsum}
\usepackage{xcolor}
\usepackage{pdfpages}
\usepackage{listings}
\definecolor{mauve}{rgb}{0.58,0,0.82}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\lstdefinestyle{DOS}
{
backgroundcolor=\color{black},
basicstyle=\color{white}\ttfamily,
breaklines=true,
postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space}
}

\lstdefinelanguage{Cypher}
{
morekeywords={MATCH, WHERE, RETURN, CREATE, ORDER, BY, ITERATE, LOOP, ON, COLLECT, AS, AND, OR, DISTINCT,
ASC, DESC, LIMIT, SKIP, UNION, UNION, ALL, COUNT, FOREACH, WITH, NOT, count, collect, sum,
min, max, avg, id, IN, labels, exists},
morestring=[b]",
basicstyle={\small\ttfamily},
keywordstyle=\color{blue},
commentstyle=\color{dkgreen},
stringstyle=\color{mauve}\ttfamily,
comment=[l]//,
numbers=left,
xleftmargin=1em,
frame=single,
framexleftmargin=2em
}

\begin{document}
\title{\textsf{Cyp2SQL -- Documentation}}

\author{Oliver Crawford\\ojc37@cam.ac.uk}

\maketitle

\abstract{\emph{Cyp2SQL} is a tool for translating graph schemas and the graph query language Cypher to both
a different database schema, and different database language (the original aim of this project was to translate
Cypher to SQL).}

\tableofcontents
\addtocontents{toc}{%
\protect\setlength{\columnsep}{5pc}
\protect\begin{multicols}{2}}

\section{Overview}

\begin{figure}[h]
\centerline{\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{toolchainVec.eps}}
\caption{The five stages of the pipeline.}
\label{toolchain}
\end{figure}

The pipeline of the tool consists of 5 stages (as shown in Figure \ref{toolchain}):

\begin{itemize}
\item A \emph{Schema Conversion} module for creating the initial database in an engine other than Neo4j (such as a relational schema in Postgres)
\item A \emph{Parsing Cypher} module, which is the first part of the translation to another database language (such as SQL). This module relies on the ANTLR framework, and other hand-written parsing code
\item An \emph{Intermediate Representation} is then built in Java
\item A \emph{Translation Module} takes this intermediate representation, and converts it to the target database language (the tool was originally designed with SQL in mind)
\item The \emph{Output Results \& Information} module executes both the original Cypher on Neo4j, and the newly created query on the other chosen database(s). The execution times of the queries may be recorded, and the results of the queries may be written to a local file if requested. In the latter case, the tool can also verify that both the number of results and the results themselves match (this is useful for debugging).
\end{itemize}

\medskip

The codebase is entirely Java. There are some additional features provided through the ANTLR tool, as described in \S \ref{sec:antlr}. There are some script/.bat files that help the tool execute SQL on Postgres and pipe the results back into the tool.

\subsection{Quick usage}
\begin{verbatim}
<-schema|-translate|-s|-t> <propertiesFile> <targetDatabaseName> <-dp|-dn|-r>
\end{verbatim}

\begin{center}
\begin{tabular}{ p{3cm} p{10.5cm} }
-s \textbar{} -schema & Runs the tool for schema conversion. \\
-t \textbar{} -translate & Runs the tool for query translation. \\
propertiesFile & Path of the properties file \texttt{c2s\_props.properties}. See \S \ref{ss:prop}. \\
targetDatabaseName & Name of the relational database for either the converted graph schema to be stored into, or for the translated queries to execute on. \\
dp \textbar{} -dn \textbar{} -r & \textbf{Only needed when in query translation mode}. See \S \ref{sec:tranQ}.
\end{tabular}
\end{center}

\section{Getting started}
The code for the tool can be found in the \emph{cyp2sql-next} repository: \url{https://gitlab.dtg.cl.cam.ac.uk/fresco-projects/cyp2sql-next/}.

\medskip

There is an initial setup procedure involved before the tool can be run. View the \texttt{README.md} for additional information.

\medskip

\textcolor{red}{NOTE: Javadoc documentation is also available through the repo.}

\medskip

\textcolor{red}{NOTE: a .JAR file is included within the repo that can be run from the command line.}

\subsection{Prerequisites}
\begin{itemize}
\item Java 1.8 or above
\item Neo4j, any version should suffice
\item \textbf{An existing graph database} in Neo4j
\item A target database engine for the queries to be translated for, such as Postgres
\item Windows or Unix OS
\item For large databases, a relatively powerful machine and available disk space (additionally, as much RAM as possible)
\end{itemize}

\subsection{File system preparation}
The easiest way to use the tool is to create a new folder, and to also create/add the following folders/files:

\begin{itemize}
\item A folder for workspace usage
\begin{itemize}
\item Metafiles created during the schema conversion will be stored here
\end{itemize}
\item A blank text file for storing the keys of the graph that may contain lists -- see \S \ref{sssec:lists}
\item A folder for storing the results from both the Neo4j database, and the target database. For example, create a folder called \texttt{results}, and then two files within this folder called \texttt{results\textbackslash neo4jResults.txt} and \texttt{results\textbackslash postgresResults.txt}
\item A new Postgres database should also be created -- in Unix this can be done via the command \texttt{createdb <nameOfNewDatabase>}
\end{itemize}

\subsection{Generating dump files from Neo4j}
The tool builds up an equivalent schema in the target database by parsing a text file that is generated via the Neo4j console.

\medskip

\textcolor{red}{NOTE: the console/Java application offered by Neo4j is currently being deprecated in favour of the new `cypher-shell'.}

\medskip

\textcolor{red}{NOTE: the dump file may be very large for large graph databases.}

\subsubsection{Windows OS}
\begin{lstlisting}[style=DOS]
java -classpath "C:\Program Files\Neo4j CE 3.0.6\bin\neo4j-desktop-3.2.2.jar" org.neo4j.shell.StartClient -path "C:\Users\ocraw\Documents\CL Internship\Neo4j graphs\Test1KOPUS" -c dump > "C:\Users\ocraw\Documents\CL Internship\Graph Dumps\dumpOPUS1k.txt"
\end{lstlisting}

\subsubsection{Unix OS}
\begin{lstlisting}[language=bash, breaklines=true, postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space}, backgroundcolor=\color{gray}, basicstyle=\color{white}\ttfamily]
neo4j-shell -path ../data/databases/dump_prov1000.graphdb/ -c dump > 2017-08-01-dump.txt
\end{lstlisting}

\subsubsection{Dump file explanation}
The dump file should look something like:

\begin{verbatim}
begin
commit
begin
create (_0:`Local` {`mono_time`:"10", `name`:"omega", `node_id`:1, `ref_count`:1, ...
create (_1:`Global` {`name`:["nginx"], `node_id`:2, `sys_time`:-782642514, `type`:2})
create (_2:`Global` {`name`:["nginx"], `node_id`:3, `sys_time`:-782642514, `type`:2})
\end{verbatim}

The tool will remove the unnecessary lines from the file, leaving two distinct types remaining to parse into a relational schema: nodes and edges (relationships).

\subsection{Configuring the properties}
\label{ss:prop}
The file \texttt{c2s\_props.properties} contains all the necessary parameters to allow the tool to function correctly. This file should be edited before running the tool.

\begin{center}
\begin{tabular}{ p{3cm} p{10.5cm} }
neo4jSchema & Path of the dump file generated from the existing Neo4j graph. \\
workspaceLocation & Path to workspace used by the tool. \\
neo4jResults & Path for the results from the Neo4j database to be stored into, when a Cypher query is executed. \\
sqlResults & Path for the results from the relational database to be stored into, when an SQL query is executed. \\
listsLocation & Path to the text file containing the list of fields that may contain lists. \\
neo4jUser & Neo4j graph database username. \\
neoPW & Neo4j graph database password. \\
postgresUser & Postgres database username. \\
postgresPW & Postgres database password. \\
\end{tabular}
\end{center}

\section{Schema conversion}
The first step of the pipeline in translating Cypher to SQL is to convert the Neo4j graph database to a relational database. Using the dump file generated from the Neo4j graph, multiple relations can be built up and executed on a relational database backend. During this process, other metafiles are also created -- these are used throughout the translation process later for completeness, and in some cases, optimisations.

\subsection{Converting the graph database to a relational schema}
For the tool to run successfully and without error, make sure all the properties in the properties file are correct. Next, a blank relational database needs to be created. The tool currently will not overwrite existing databases -- if an error has occurred when writing to the database, it must be wiped clean before the tool can run again.

\medskip

\begin{verbatim}
-s C:\Users\ocraw\IdeaProjects\Cypher_SQL_Translation\c2s_props.properties testDB
\end{verbatim}

\medskip

\begin{itemize}
\item The first argument (\texttt{-s}) signals to the program that it should be performing the schema conversion
\item The second argument is the location of the properties file
\item The third argument is the name of the recently created blank database
\end{itemize}

\medskip

If the tool runs successfully and without error, the following should have happened:

\begin{itemize}
\item The recently created database will now be populated with tables and data -- see \S \ref{ssec:grreltheory}
\item The workspace folder will now contain four metafiles -- see \S \ref{ssec:meta}
\end{itemize}

\medskip

\textcolor{red}{NOTE: deleting the workspace folder and its metafiles will cause some translations to fail. If this is done accidentally, the whole graph schema must be converted again from scratch.}

\subsection{Graph to relational: theory}
\label{ssec:grreltheory}
All the nodes from the graph are stored in one relation (nodes), and all the relationships are stored in a separate relation (edges).

There are some optimisations and additional implementation aspects. Storing all the attributes of the nodes in one relation, where there are multiple labels, will lead to a lot of NULL values populating the relation. To combat this, each individual label also has its own relation. The translator tool decides at run time the optimal relation to join.

Furthermore, each relationship type has its own relation. Although this optimisation is quicker, the trade-off will be roughly double the amount of disk space required to store all the relations, as well as the additional complexity in the translation tool.

For graph functions to be computable on the relational side, the schema converter also builds two materialised views –- \texttt{adjlist\_from} \& \texttt{adjlist\_to}.

Some other functions are also committed to the database at this point:

\begin{center}
\begin{tabular}{ p{4.5cm} p{9cm} }
\texttt{array\_unique(arr anyarray)} & Helper function for the \texttt{cypher\_iterate} function (see below). This function takes an array of any type as input, and returns an array of the same type, but with duplicate elements removed. \\

\texttt{cypher\_iterate(integer[])} & Function for performing the iterate function designed for this project. \\

\texttt{doforeachfunc(integer[], field TEXT, newv TEXT)} & Function to help perform the semantics of Cypher’s \texttt{FOREACH} keyword.
\end{tabular}
\end{center}

\medskip

\textcolor{red}{NOTE: the class \texttt{PostgresConstants.java} contains the list of additional materialised views and functions committed to the database during the schema conversion process.}

\medskip

Figure \ref{schemaConv} shows in detail an example conversion (the example is based on an OPUS graph with 100{,}000 nodes).

\begin{figure}[p]
\centerline{\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{schemaConv.eps}}
\caption{Schema conversion outline.}
\label{schemaConv}
\end{figure}

\subsection{Metafiles created during schema conversion}
\label{ssec:meta}
\texttt{meta\_nodeProps.txt} -- \textbf{Creation:} all properties of the nodes are stored in this file. \textbf{Usage:} currently two major uses. One is to help the Neo4j database driver to return the correct results to the user (for example, when a wildcard is used in the query). Secondly, it is used in the cases where a GROUP BY is required in the SQL statement to be executed.

\bigskip

\texttt{meta\_labelProps.txt} -- \textbf{Creation:} a slightly more detailed version of the file above. The file contains the list of properties of all the nodes, but is sectioned into the individual label types of the nodes.

Example:
\begin{verbatim}
*country*
name
population
dialling_code
*city*
name
population
state
ave_house_price
\end{verbatim}

\textbf{Usage:} the file is used to see if any possible optimisations to the SQL query are achievable. This is done by finding unique nodes for a property. In the example above, if the property/key `state' only belongs to the label `city', then the tool can use this information to its advantage.

See the method \texttt{getLabelMapping()} in \texttt{C2SMain.java} for further information. Information from this file is stored in the \texttt{labelProps} map (publicly accessible using \texttt{C2SMain.labelProps}).

\bigskip

\texttt{meta\_rels.txt} -- \textbf{Creation:} all the types of edges in the Neo4j graph are stored in this file.
\textbf{Usage:} currently only used in the case where the tool can be used to delete information from the database (this feature has not been thoroughly tested yet though, so caution is advised if a query which is being used to delete data is to be executed).

See the method \texttt{getLabelMapping()} in \texttt{C2SMain.java} for further information. Information from this file is stored in the \texttt{allRelTypes} map (publicly accessible using \texttt{C2SMain.allRelTypes}).

\bigskip

\texttt{meta\_labelNames.txt} -- \textbf{Creation:} all the names of the labels in the Neo4j graph are stored.
\textbf{Usage:} another optimisation technique -- helps the tool choose the best relation to use to complete the query correctly.

\section{Translation of queries}
\label{sec:tranQ}
\subsection{Usage}
The tool requires a running Neo4j instance to be up and running.

\medskip

\begin{verbatim}
-t \..\..\c2s_props.properties testDB <-dp|-dn|-r>
\end{verbatim}

\medskip

\begin{itemize}
\item The first argument (\texttt{-t}) signals to the program that it should be performing the translation of queries
\item The second argument is the location of the properties file
\item The third argument is the name of the relational database that the queries are to be executed on
\item The fourth argument is another flag that gives the user some choice in what they want to do -- see the table below for further guidance
\end{itemize}

\begin{center}
\begin{tabular}{ p{2cm} p{11.5cm} }
\texttt{-dp or -dn} & \textbf{Debug interface}. This function allows the user to input queries through the command line interface for translation. Useful for debugging. If \texttt{-dp} is used, the results from both the databases are printed to local files. If \texttt{-dn} is set, this does not occur. \\

\texttt{-r} & \textbf{Run as normal}. This function allows the user at the command line to input queries for translation. Unlike \texttt{-dp} or \texttt{-dn}, this will not execute Cypher on Neo4j, and will only output the results from the target database engine.
\end{tabular}
\end{center}

\medskip

\textcolor{red}{NOTE: the script to run Postgres on Unix systems may not have the right permissions to run with the tool. \texttt{chmod +x auto\_run\_pg.sh} should fix the issue.}

\medskip

The user at this point should be presented with a command line interface where Cypher can be typed into:

\begin{figure}[h]
\centerline{\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{ss1.png}}
\caption{Screenshot of debug mode.}
\label{ss1}
\end{figure}

\begin{figure}[h]
\centerline{\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{ss2.png}}
\caption{Screenshot of run mode.}
\label{ss2}
\end{figure}

\subsection{Query translation}
The query is first tokenised using the ANTLR generated classes for the Cypher language. See \S \ref{sec:antlr} for more details.

An intermediate representation in the form of a DecodedQuery object is the built. This object is then referred to when translating to the target language (such as SQL). See \S \ref{ssec:workflow} for more detailed descriptions.

Multiple different classes can be used to perform the translation to SQL. Likewise, if a new type of translation is needed, the \texttt{AbstractConversion.java} and \texttt{AbstractTranslation.java} classes can be extended to suit.

\subsubsection{Dealing with Cypher lists}
\label{sssec:lists}
Lists in Cypher present one current challenge for the translation tool. Although the translation can be done, there is no current mechanism for detecting cleanly which keys of nodes/relationships may have type list.

To solve this problem (and to possibly fix any issues encountered with lists), a blank file should be created (with a name such as \texttt{lists.txt}). This file should then be referred to in the \texttt{c2s\_ props.properties} file (see \S \ref{ss:prop}).

The file should contain a list, separated by new lines, of each possible key in the graph that may contain a list. This will aid the translator in producing the correct result.

\subsubsection{Where clauses}
Where clauses are converted into properties of nodes/relationships. For example, if there is a Cypher clause, \texttt{MATCH (n) WHERE n.a = 1 AND n.b = 2 RETURN n}, then the translator tool associates the two predicates to the node \texttt{n}. When the translation to SQL occurs, the properties of node \texttt{n} are then extracted and converted.

The whole where clause is decoded in \texttt{CypherTranslator.java}. The initial clause is separated into its individual parts (they will be separated by either AND/OR). A \texttt{CypWhere} object is then created for each component, to store:

\begin{itemize}
\item The position it appears in the clause (the indexing counting left to right)
\item Any brackets that appear before or after the clause (to ensure consistent ordering later)
\item The boolean operator that immediately follows it
\end{itemize}

Depending on what the clause itself is conveying, it is parsed accordingly. When it has been parsed, the properties and semantics need to be stored. As mentioned above, they are stored with the nodes/relationships that they belong with. For example, if the where clause says something like \texttt{a.name = `me'}, the properties and semantics of the clause will be stored with the node that has the id \texttt{`a'} (assuming \texttt{`a'} is of course a node and not a relationship).

The format of how it is stored with the nodes/relationships is initially quite confusing (and in all honesty a bit silly, could do with a refactor), but can be quickly described with this guide:

\bigskip

\begin{lstlisting}[language=Cypher]
MATCH p=shortestPath((f {name:"omega"})-[*1..6]->(t:Meta))
RETURN count(t);
\end{lstlisting}

Property stored with node: \texttt{[name="omega"]}

In this example, as the property is described within the node itself, the property is simple to understand and parse.

\bigskip

\begin{lstlisting}[language=Cypher]
MATCH (a:Meta)
WHERE a.sys_time < 0
RETURN count(a);
\end{lstlisting}

Property stored with node: \texttt{[sys\_ time="p1null@null\$lt\#0\#tl"]}

\begin{itemize}
\item Already, the syntax looks very different, but can be broken down easily
\item \texttt{p1} denotes that the component \texttt{a.sys\_time < 0} is in the first position of the clause (in fact, in this example, it is the only component)
\item The \texttt{null} before the @ symbol means there is no boolean operator (such as AND or OR) that follows this component
\item The \texttt{null} after the @ symbol means that the component is not associated with any bracketing
\item The value after the \$ sign, \texttt{lt\#0\#tl} describes the predicate itself:
\begin{itemize}
\item \texttt{lt} is the code for the \texttt{<} symbol
\item The operand 0 is contained within the \# symbols
\end{itemize}
\end{itemize}

\bigskip
\newpage

\begin{lstlisting}[language=Cypher]
MATCH (a)
WHERE a.node_id < 345
OR (
(a.node_id > 800 AND 'Process' in labels(a))
OR a.node_id = 983
)
RETURN count(a);
\end{lstlisting}

Property stored with node:
\texttt{[node\_id="p1or@null\$lt\#345\#tl\textasciitilde p2and@((\$gt\#800\#tg\textasciitilde p4null@)\$eq\#983\#qe", label="p3or@)\$eq\#process\#qe"]}

\begin{itemize}
\item In this case, two fields of the node have properties associated with them
\item When multiple properties apply to one field, the \textasciitilde {} symbol separates the individual components that will be translated
\item When brackets are involved (notice in the original Cypher input the 2 left brackets before \texttt{a.node\_id > 800}), these are stored after the @ symbol (in their literal form)
\begin{itemize}
\item Extra care must be taken in the code in the case of functions such as \texttt{labels(a)} at the end of a component (only one bracket should be recorded and not two)
\end{itemize}
\end{itemize}

\begin{center}
\begin{tabular}{ p{2.5cm} p{11cm} }
\texttt{eq\#...\#qe} & Equality: \texttt{WHERE idA.property = someValue}. \\
\texttt{ne\#...\#en} & Not equal to: \texttt{WHERE idA.property <> someValue}. \\
\texttt{lt\#...\#tl} & Less than: \texttt{WHERE idA.property < someValue}. \\
\texttt{gt\#...\#tg} & Greater than: \texttt{WHERE idA.property > someValue}. \\
\texttt{le\#...\#el} & Less than or equal to: \texttt{WHERE idA.property <= someValue}. \\
\texttt{ge\#...\#eg} & Greater than or equal to: \texttt{WHERE idA.property >= someValue}. \\
\texttt{ex\#...\#xe} & Exists function: \texttt{WHERE exists(idA.property)}. \\
\texttt{nx\#...\#xn} & Exists function (not): \texttt{WHERE not exists(idA.property = someValue)}. \\
\texttt{in\#...\#ni} & IN: \texttt{WHERE `someLabel' in labels(idA)}. \\
\texttt{anyin\#...\#niyna} & Any (where IN is also used): \texttt{WHERE any(var IN idA.property WHERE var IN [someList]}. \\
\texttt{anyeq\#...\#qeyna} & Any (where IN is also used): \texttt{WHERE any(var IN idA.property WHERE var = someValue}. \\
\texttt{isn\#...\#nsi} & IS NULL: \texttt{WHERE idA.proeprty IS NULL}. \\
\texttt{non\#...\#non} & IS NOT NULL: \texttt{WHERE idA.proeprty IS NOT NULL}.
\end{tabular}
\end{center}

\subsubsection{Iterate semantics}
\label{sssec:iterate}
As part of an extension to my original dissertation on this tool, additional functionality to the graph model was included. The additional feature is looking at querying repeating patterns within a graph. A user defines a pattern they wish to search, and a schema for how to perform the iterative matching. The query will store all the nodes “touched” by the query, returning them similarly to traditional Cypher queries.

To make the extension simple, it was important to define the syntax in a clear and distinct manner, as demonstrated below:

\begin{lstlisting}[language=Cypher]
ITERATE MATCH (a {host: `google'})-->(b:Website)
-->(c {domain: `net'})
LOOP c ON b COLLECT n
RETURN n.host
\end{lstlisting}

The \texttt{ITERATE} keyword makes it both clear that the Cypher will attempt to iterate over a graph, and it is used as a flag for the tool, so that it can expect to parse the query correctly. The remaining syntax encompasses the iterate pattern: \texttt{LOOP (1) ON (2) COLLECT (3)}.

\texttt{(1)} is the node ID holding the results of intermediate queries. \texttt{(2)} marks the starting point in the query for the previous results to start from. \texttt{(3)} is an extra identifier that is separate from the main query, and is what is used in the \texttt{RETURN} clause.

To demonstrate what this new syntax can do, consider the graph in Figure \ref{iterGraphFig}.

\begin{figure}[h]
\centerline{\includegraphics[scale=0.7]{iterateGraph.png}}
\caption{A trivial graph consisting of thirteen nodes with three different labels -- blue for `l1', green for `l2', and red for `l3'.}
\label{iterGraphFig}
\end{figure}

\newpage
Figure \ref{iterGraphFig} shows a common pattern, \texttt{BLUE --> GREEN --> RED --> BLUE}. Finding all the blue nodes at the end of this pattern \emph{iteratively} would be hard to do with the existing Cypher syntax in a concise and clear manner. It is a more tractable problem with the \texttt{ITERATE} syntax:

\begin{lstlisting}[language = Cypher]
ITERATE MATCH (a:l1)-->(b:l2)-->(c:l3)-->(d:l1)
LOOP d ON a COLLECT n
RETURN n.uid;
\end{lstlisting}

Executing this query on Postgres returns the results shown in Table \ref{table:queryIter}.

\begin{table}[!h]
\caption{Results of Iterate Query.} % title of Table
\centering % used for centering table
\begin{tabular}{c} % centered columns (4 columns)
\hline\hline %inserts double horizontal lines
n.uid \\ [0.5ex] % inserts table
%heading
\hline % inserts single horizontal line
9 \\
9 \\
9 \\
10 \\
13 \\
13 \\
13 \\
13 \\ [1ex] % [1ex] adds vertical space
\hline %inserts single line
\end{tabular}
\label{table:queryIter} % is used to refer this table in the text
\end{table}

In comparison, running just the \texttt{MATCH} query on Neo4j once will return only six results. The extra two results that the \texttt{ITERATE} extension finds comes from the fact that the nodes with ID `9' and `10' are being used in another iteration of the \texttt{MATCH} clause. Hence the node with ID `13' is returned a further two times.

\subsection{Workflow through the code}
\label{ssec:workflow}
On the following pages is an example of how a query is translated from Cypher to SQL through the tool. Included with the flow are descriptions to help the reader.

\newpage
\includepdf[pages={-}]{workflow.pdf}

\section{ANTLR framework}
\label{sec:antlr}
The current tool uses ANTLR v4\footnote{\url{http://www.antlr.org/download.html}} as a way of parsing the initial Cypher query, both into tokens and as a more effective way of parsing information out of the query. The ANTLR framework requires a grammar of Cypher to work, and this has been obtained from the \emph{openCypher} project\footnote{\url{https://s3.amazonaws.com/artifacts.opencypher.org/M06/Cypher.g4}}.

The guide for setting up the parsers and lexers has been adapted from the ANTLR v4 `Getting Started' guide\footnote{\url{https://github.com/antlr/antlr4/blob/master/doc/getting-started.md}}.

\medskip

\textcolor{red}{NOTE: the parser and lexer should already be working and configured within the \emph{Cyp2SQL} code itself. The information below would be useful in the case of an update to either the Cypher grammar file, or to the ANTLR library. See the package \texttt{parsing\_lexing} for the existing configuration.}

\begin{figure}[h]
\centerline{\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{ss3.png}}
\caption{Parse tree of a Cypher query with ANTLR.}
\label{antlrs}
\end{figure}

\subsection{ANTLR setting up}

Firstly, a new temporary folder should be made -- add to it the latest \texttt{Cypher.g4} grammar file.\footnote{One of the keywords in the original file was named \texttt{return}, but this clashes with the tool, and so it has been renamed to \texttt{returnX}} Then run the following command:

\begin{lstlisting}[style=DOS]
java -jar "C:\Program Files\Java\jdk1.8.0_101\lib\antlr-4.7-complete.jar" Cypher.g4
\end{lstlisting}

Followed by this command:

\begin{lstlisting}[style=DOS]
javac -classpath .;"C:\Program Files\Java\jdk1.8.0_101\lib\antlr-4.7-complete.jar" Cypher*.java
\end{lstlisting}

It is a good idea to test the installation at this point:

\begin{lstlisting}[style=DOS]
java -cp ".;C:\Program Files\Java\jdk1.8.0_101\lib\antlr-4.7-complete.jar" org.antlr.v4.gui.TestRig Cypher cypher -gui
\end{lstlisting}

Assuming a valid installation, commands can now be typed into the console - \texttt{\^{}Z} will end the input and hopefully produce an image, such as the one in Figure \ref{antlrs}.

The source code files generated by the tool are contained within the package \texttt{parsing\_lexing}.

\section{Current Cypher translation list}
The tool cannot currently translate the complete set of Cypher available to use. Below is a detailed guide of what can be translated. Further translations are more than possible by viewing/adapting the source code to suit.

\medskip

\textcolor{red}{NOTE: information gathered below has been adapted from the full list of Cypher commands obtained from the Neo4j Cypher reference card: \url{https://neo4j.com/docs/cypher-refcard/current/}}

\medskip

\textcolor{red}{NOTE: quantifiers from regular expressions may be used below to make the list more concise. As a quick reference, ${*}$ = zero or more times; + = one or more times; ? = zero or one time.}

\medskip

\subsection*{\texttt{MATCH ... RETURN ...}}
The queries below will be translated via the \texttt{NoRels.java} class.

\medskip

\begin{lstlisting}[language = Cypher]
MATCH (idA)
RETURN [DISTINCT]? idA

MATCH (idA:labelA)
RETURN idA.propertyA

MATCH (idA:labelA:labelB {keyA:valueA})
RETURN idA.propertyA AS alias_value, idA.propertyB AS other_alias

MATCH (idA)
RETURN [count(idA)|collect(idA.propertyA)|sum(idA.propertyA)|
min(idA.propertyA)|max(idA.propertyA)|avg(idA.propertyA)]
\end{lstlisting}

\subsection*{\texttt{ORDER BY, SKIP, and LIMIT}}
Queries containing these keywords need only a small adjustment to translate.

\medskip

\begin{lstlisting}[language = Cypher]
MATCH (idA)
RETURN idA.propertyA
ORDER BY idA.propertyB [ASC|DESC]?
[SKIP skipValue]? [LIMIT limitValue]?
\end{lstlisting}

\subsection*{\texttt{MATCH ... WHERE ... RETURN}}
These are also translated to SQL in the class \texttt{NoRels.java}.

\medskip

\begin{lstlisting}[language = Cypher]
MATCH (idA)
WHERE [NOT]? idA.propertyA [=|<>|<|>|<=|>=] value
RETURN count(idA)

MATCH (idA)
WHERE idA.propertyA < valueA [AND|OR] idB.propertyB > valueB
RETURN count(idA)
\end{lstlisting}

\subsection*{Paths -- \texttt{MATCH ()-->()-->() ...}}
These are translated to SQL in the class \texttt{MultipleRel.java}.

\medskip

\begin{lstlisting}[language = Cypher]
MATCH (idA)-->(idB)-->(idC)-->(idD)
WHERE id(idD) < some_value
RETURN count(idA) AS some_alias

MATCH (idA)<-[[relID]?:REL_TYPE]-(idB)
RETURN idB
ORDER BY idB.propertyA ASC

MATCH ()-[r:REL_TYPE {keyA:valueA}]-()
RETURN r.propertyA
\end{lstlisting}

\subsection*{With -- \texttt{... WITH ... RETURN}}
These are translated to SQL in the class \texttt{WithSQL.java}, although additional parsing of the Cypher input will also occur in either the class \texttt{With\_Cypher.java} or \texttt{Multiple\_With\_Cypher.java}.

\medskip

\begin{lstlisting}[language = Cypher]
MATCH (idA:labelA)-[rel]->(idB:labelB)
WITH idA, count(rel) AS alias
WHERE alias > value
RETURN idA.propertyA, alias
ORDER BY alias DESC

MATCH (idA)
WITH idA
ORDER BY idA.propertyA DESC
LIMIT 3
RETURN collect(idA.propertyA)

MATCH (idA)
WHERE someLabel IN labels(idA)
WITH idA
MATCH (idB)
WHERE idB.propertyB = idA.propertyA
RETURN count(idA)
\end{lstlisting}

\subsection*{\texttt{UNION [ALL]?}}
The individual components of the input are dealt with normally, and then combined together with the correct \texttt{UNION} semantics in the class \texttt{UnionSQL.java}.

\medskip

\begin{lstlisting}[language = Cypher]
MATCH (idA)-[:rel1]->(idB)
RETURN idB.propertyA
[UNION|UNION ALL]
MATCH (idA)-[:rel2]->(idB)
RETURN idB.propertyA
\end{lstlisting}

\newpage

\subsection*{Variable length paths -- \texttt{-[*x..y]->}}
This subset of Cypher is dealt with using the \texttt{SingleVar.java} class.

\medskip

\begin{lstlisting}[language = Cypher]
MATCH (idA:labelA)-[*x..y]->(idB:labelB)
RETURN count(idB)

//x and y represent integer values for the length of the path
//being traversed. These queries, unlike the paths shown earlier,
//only currently work for one pattern - it cannot be extended
//in the following way:

MATCH (idA:labelA)-[*x..y]->(idB:labelB)-[*x2..y2]->(idC)
RETURN count(idC)

//furthermore, a direction must be specified.
\end{lstlisting}

\subsection*{Shortest path queries}
This subset of Cypher is dealt with using the \texttt{ShortestPath.java} class.

\medskip

\begin{lstlisting}[language = Cypher]
MATCH p=shortestPath((idA {key:value})-[*1..x]->(idB:labelB))
RETURN count(idB)
\end{lstlisting}

\subsection*{Functions -- \texttt{exists(), labels(), id()}}
These keywords are parsed in the class \texttt{CypherTranslator.java}, where they are stored with the node/relationship that they are referring to.

\medskip

\begin{lstlisting}[language = Cypher]
MATCH (idA)
WHERE exists(idA.propertyA) AND exists(idA.propertyB)
RETURN count(idA)

MATCH (idA)-[rel]->(idB)
WHERE NOT id(idB) = value OR "someLabel" IN labels(idB)
RETURN idA.propertyA
\end{lstlisting}

\subsection*{\texttt{... WHERE ... IN} and \texttt{any()}}
These keywords are parsed in the class \texttt{CypherTranslator.java}, where they are stored with the node/relationship that they are referring to.

\medskip

\begin{lstlisting}[language = Cypher]
MATCH (idA)
WHERE id(idA) IN [val1, val2, ..., valN]
RETURN idA.propertyA

MATCH (idA)
WHERE any(someV IN labels(idA) WHERE someV IN [val1, ..., valN])
RETURN count(idA)

MATCH (idA)
WHERE any(someVar IN idA.propertyA WHERE someVar = value)
RETURN count(idA)
\end{lstlisting}

\subsection*{\texttt{ITERATE MATCH ...}}
This additional keyword is parsed and translated in the class \texttt{Iterate\_Cypher.java}. See \S \ref{sssec:iterate} for more details.

\medskip

\begin{lstlisting}[language = Cypher]
ITERATE MATCH (idA)-->(idB)-...->(idX)
LOOP idX ON idA COLLECT idN
RETURN idN.propertyA
\end{lstlisting}

\subsection{Known issues}
\begin{itemize}
\item The use of \texttt{WITH} is still quite constrained in its use
\item When predicates in the \texttt{WHERE} clause are added, they may sometimes lead to incorrect translations (mostly in the \texttt{WITH} parts of a multiple relationship Cypher input). It can be solved in some cases by adding the predicates in a different way in the original input (see example below)
\end{itemize}

\begin{lstlisting}[language = Cypher]
MATCH (a)-[e]->(b)-[f]->(c)
WHERE a.type = b.type AND c.pid < b.pid
RETURN count(f);

// instead of

MATCH (a)-[e]->(b)-[f]->(c)
WHERE a.type = b.type AND b.pid > c.pid
RETURN count(f);

// where 'b' now has 2 properties associated with it,
// and will not work (currently).
\end{lstlisting}

\subsection{List of test queries from OPUS}
\begin{verbatim}
MATCH p=shortestPath((f {name:"omega"})-[*1..6]->(t:Meta)) RETURN count(t);

MATCH (a)-[*1..3]->(c:Process) RETURN count(c);

MATCH (a:Local)-[*4..9]->(b)
RETURN DISTINCT b.node_id, b.sys_time AS time_alias ORDER BY b.node_id DESC;

MATCH (n {name:["/var/db/entropy/saved-entropy.7", "/var/db/entropy/saved-entropy.8"]})
RETURN n.node_id ORDER BY n.node_id ASC;

MATCH (a:Global)-[]->(b) RETURN b.node_id AS conn_id;

MATCH (a:Global)-[]->(b) RETURN count(b.node_id);

MATCH (a:Global)-->(b:Local)-->(c:Process)<--(d:Local)<--(b) RETURN count(b);

MATCH ()-[r]-() RETURN DISTINCT r.state ORDER BY r.state;

MATCH (n:Local)<--(m:Global)
RETURN m.node_id AS thing, m.type AS ty ORDER BY m.sys_time LIMIT 3;

MATCH (a:Global) RETURN count(a) AS funky;

MATCH (a:Meta) WHERE a.sys_time < 0 RETURN count(a);

MATCH (a:Meta) WHERE a.sys_time < 0 OR a.node_id > 845 RETURN count(a);

MATCH (a)-[r:LOC_OBJ]-(b) RETURN b.name, r.state ORDER BY b.node_id ASC LIMIT 15;

MATCH ()<-[r:LOC_OBJ {state:12}]-(idA {type:2}) RETURN count(r);

MATCH (n) WHERE id(n) = 345 RETURN n.mono_time, n.sys_time, n.name;

MATCH (a)-->(b)-->(c)-->(d) WHERE id(d) < 123 RETURN count(a) AS cool;

MATCH (n) WHERE exists(n.value) AND exists(n.timestamp) RETURN count(n);

MATCH (n)--()--()--()--(n) WHERE exists(n.status) RETURN count(n);

MATCH (s)-[e]-(d)
WHERE id(s) = 349 AND NOT 'Process' in labels(s) AND NOT 'Global' in labels(d)
RETURN d.node_id ORDER BY d.node_id ASC;

MATCH (n:Process)<-[e:PROC_OBJ]-(c:Local)
WHERE id(n) = 916 AND e.state in [5]
RETURN c.name, e.state ORDER BY c.name DESC;

MATCH (a)-[e]-(b)
WHERE id(a) IN [100, 200, 300, 400] AND id(b) IN [101, 201, 202, 302, 404]
RETURN e.state;

MATCH (a) WHERE any(name in a.name WHERE name = 'uid') RETURN count(a);

MATCH (a) WHERE any(lab in labels(a) WHERE lab IN ['Global', 'Meta'])
RETURN count(a);

MATCH (n) WHERE 'Process' in labels(n)
WITH n
MATCH (m) WHERE m.status = n.status RETURN count(n);

MATCH (n) WHERE 'Local' in labels(n) AND NOT exists(n.pid)
WITH n
MATCH (m:Global)-[r]->(n) WHERE id(m) > 900 RETURN n.node_id, r.state;

MATCH (n:Global)-->(m:Local) WHERE n.node_id < m.node_id RETURN count(m);

MATCH (n:Meta)<--(m:Process)-->(p)
WHERE n.node_id > m.node_id AND p.node_id <= m.node_id RETURN count(m);

MATCH (n) WHERE id(n) < 3
WITH n
MATCH (m) WHERE id(m) < id(n)
WITH m
MATCH (p) WHERE p.node_id < m.node_id RETURN count(p);

MATCH (n) WHERE 'Meta' in labels(n) OR any(name in n.name WHERE name = 'postgres')
WITH n
MATCH (m:Process) WHERE id(m) > id(n)
WITH m
MATCH (p)-->(m)
WITH p
MATCH (j)<-[:PROC_OBJ_PREV]-(p) WHERE p.sys_time = j.sys_time RETURN count(j);

MATCH (a:Global {name:'postgres'})-->(b:Global)
WITH b
MATCH (c) WHERE c.sys_time = b.sys_time
WITH c
MATCH (c)<--(d) RETURN DISTINCT d.node_id ORDER BY d.node_id LIMIT 5;

MATCH (a:Meta) RETURN count(distinct a.name);

MATCH (a:Local)-->(b)<--(c:Process)<--(d) RETURN min(d.node_id);

MATCH (n)
WHERE 'Global' in labels(n) AND any(name in n.name WHERE name = 'master')
OR (exists(n.pid) AND n.status = 2)
WITH n
MATCH (m:Meta) WHERE m.node_id > n.node_id RETURN DISTINCT n LIMIT 10;

MATCH (a)
WHERE a.node_id < 345
OR ((a.node_id > 800 AND 'Process' in labels(a)) OR a.node_id = 983)
RETURN count(a);

MATCH (a)
WHERE (any(x in a.name where x = 'master')
OR any(y in a.value where y in ['postgres', 'nginx']))
AND ('Global' in labels(a) OR 'Meta' in labels(a))
RETURN count(a);

MATCH (a)-[e]->(b)-[f]->(c)
WHERE a.type = b.type AND c.pid < b.pid RETURN count(f);

MATCH (a)-[z]->(b)-[w]->(c)
WHERE a.node_id < b.node_id RETURN w.state, c.type ORDER BY c.node_id ASC;

MATCH (a)-[e]->(b:Process) WHERE e.state > 5
WITH b
MATCH (c) WHERE (exists(c.pid) AND c.pid < b.pid)
WITH c
MATCH (c)<--(d:Local)
WHERE any(n in d.name WHERE n = '4') RETURN count(d) AS cool_thing;

MATCH (a)-->(b)-->(c) WHERE c.node_id < b.node_id
WITH c
MATCH (d)--(c) WHERE exists(d.ref_count)
WITH d
MATCH (e)-->(d)<--(f) WHERE f.node_id > e.node_id
WITH f
MATCH (g)<-[ww]-(f) WHERE ww.state = 5
WITH g
MATCH (g)-->(ii)-->(i) RETURN DISTINCT i.node_id ORDER BY i.node_id ASC;

MATCH (a)-->(b)-->(c:Process)
WHERE a.node_id < b.node_id
RETURN a.node_id, b.node_id, case c.type when 3 then 'boo' else 'hiss' end

MATCH (a:Global)-[:LOC_OBJ]->(b)
WITH a, count(b) AS num_things WHERE num_things > 2
RETURN a.name ORDER BY a.sys_time DESC;
\end{verbatim}

\addtocontents{toc}{\protect\end{multicols}}
\end{document}

### 2017-08-15 src/main/java/database/Neo4jDriver.java

@@ -144,6 +144,8 @@ public class Neo4jDriver {
             return "sum";
         } else if (origKey.contains("avg(")) {
             return "avg";
+        } else if (origKey.startsWith("case")) {
+            return "case";
         } else if (origKey.contains(".")) {
             StringBuilder newKey = new StringBuilder();
             String[] parts = origKey.split("\\.");

### 2017-08-15 src/main/java/query_translation/sql/conversion_types/With_Cypher.java

@@ -15,12 +15,12 @@ public class With_Cypher extends AbstractConversion {
      */
     @Override
     public String convertQuery(String cypher) {
-        int posOfMatch = cypher.toLowerCase().indexOf("match");
+        int posOfMatch = cypher.toLowerCase().lastIndexOf("match");
         int posOfWhere = cypher.toLowerCase().indexOf("where");
         int posOfOrderBy = cypher.toLowerCase().indexOf("order by");
 
         // from the tokens in the original Cypher query, decide on the most appropriate method for translation.
-        if (posOfMatch != -1) return withMatch(cypher);
+        if (posOfMatch != -1 && posOfMatch != 0) return withMatch(cypher);
         if (posOfOrderBy == -1 || (posOfWhere != -1 && (posOfWhere < posOfOrderBy))) return withWhere(cypher);
         else return withOB(cypher);
     }

### 2017-08-15 src/main/java/query_translation/sql/utilities_sql/MultipleRel.java

@@ -206,148 +206,149 @@ public class MultipleRel extends AbstractTranslation {
             if (cR.getCaseString() != null) {
                 // if caseNode is false, then the type must be a relationship.
                 boolean caseNode = (cR.getType().equals("node"));
-                String replacement = caseNode ? "n01." + cR.getField() : "a." + cR.getField();
+                int posInClause = 1;
+                if (caseNode) posInClause = cR.getPosInClause();
+                String replacement = caseNode ? "n0" + posInClause + "." + cR.getField() : "a." + cR.getField();
                 String caseString = cR.getCaseString().replace(cR.getNodeID() + "." + cR.getField(), replacement);
-                safSQL.append(caseString).append(" ");
+                safSQL.append(caseString).append(", ");
                 if (caseNode) {
                     needNodeTable = true;
                     nodeTableCount++;
                 } else {
                     relsNeeded = TranslateUtils.addToRelsNeeded(relsNeeded, "a");
                 }
-                break;
-            }
+            } else {
+                boolean usesExistingWith = false;
 
-            boolean usesExistingWith = false;
+                for (String z : WithSQL.withMapping.keySet()) {
+                    if (cR.getNodeID().equals(z)) {
+                        usesExistingWith = true;
 
-            for (String z : WithSQL.withMapping.keySet()) {
-                if (cR.getNodeID().equals(z)) {
-                    usesExistingWith = true;
-
-                    String prop = cR.getField();
-                    if (cR.hasAggFunc()) {
-                        safSQL.append(CypAggFuncs.sqlEquiv(cR.getAggFunc()));
-                    }
-                    if (cR.getCount() > 0) {
-                        safSQL.append("count(");
-                        if (cR.getCount() == 2) safSQL.append("distinct ");
-                    }
-                    if (cR.getCaseString() != null) {
-                        String caseString = cR.getCaseString().replace(cR.getNodeID() + "." + cR.getField(),
-                                WithSQL.withMapping.get(z) + "." + cR.getField());
-                        safSQL.append(caseString).append(", ");
-                    } else {
-                        if (prop != null) {
-                            safSQL.append(WithSQL.withMapping.get(z)).append(".").append(prop)
-                                    .append(TranslateUtils.useAlias(cR.getNodeID(), cR.getField(), alias))
-                                    .append(", ");
-                        } else {
-                            safSQL.append(WithSQL.withMapping.get(z))
[…]

### 2017-08-15 src/main/java/query_translation/sql/utilities_sql/SQLTranslate.java

@@ -37,7 +37,7 @@ public class SQLTranslate {
             NoRels nr = new NoRels();
             sql = nr.translate(sql, decodedQuery);
         } else if (decodedQuery.getMc().isVarRel() && decodedQuery.getMc().getRels().size() == 1) {
-            SingleVarAdjList singleV = new SingleVarAdjList();
+            SingleVar singleV = new SingleVar();
             sql = singleV.translate(sql, decodedQuery);
         } else {
             MultipleRel mr = new MultipleRel();

### 2017-08-15 src/main/java/query_translation/sql/utilities_sql/SingleVar.java

@@ -7,7 +7,7 @@ import java.util.Map;
 /**
  * Class for translating Cypher with a single variable length path to SQL.
  */
-public class SingleVarAdjList extends AbstractTranslation {
+public class SingleVar extends AbstractTranslation {
     private static StringBuilder getFinalSelect(StringBuilder sql, ReturnClause returnC,
                                                 Map<String, String> alias, CypNode cn2,
                                                 int amountHigh, boolean usesDistinct) {

### 2017-08-15 src/main/java/query_translation/sql/utilities_sql/WithSQL.java

@@ -73,9 +73,10 @@ public class WithSQL {
     private static StringBuilder getOrderByForWith(StringBuilder sWith, ArrayList<String> tokens) {
         int posOfOrder = tokens.indexOf("order");
         sWith.append(" ORDER BY ");
+
         while (true) {
-            String field = tokens.get(posOfOrder + 2);
-            String dir = (posOfOrder + 3 >= tokens.size()) ? null : tokens.get(posOfOrder + 3);
+            String field = tokens.get(posOfOrder + 4);
+            String dir = (posOfOrder + 5 >= tokens.size()) ? null : tokens.get(posOfOrder + 5);
             if (dir != null) {
                 dir = (dir.equals("asc") || dir.equals("desc")) ? dir : null;
             }
@@ -108,6 +109,9 @@ public class WithSQL {
 
     private static StringBuilder getSelectForWith(StringBuilder sWith, ArrayList<String> tokens, DecodedQuery dQ) {
         sWith.append("SELECT ");
+
+        boolean hasDistinct = tokens.contains("distinct");
+        if (hasDistinct) tokens.remove("distinct");
         int posOfReturn = tokens.indexOf("return");
 
         for (int i = posOfReturn + 1; i < tokens.size(); i++) {
@@ -132,6 +136,8 @@ public class WithSQL {
                 }
 
                 String[] idAndProp = returnStmt.toString().split("\\.");
+
+                if (hasDistinct) sWith.append("distinct ");
 
                 boolean sWithChanged = false;
                 for (CypReturn cR : dQ.getRc().getItems()) {

### 2017-08-15 src/main/java/translator/CypherTranslator.java

@@ -611,9 +611,7 @@ class CypherTranslator {
                     caseString.append(s).append(" ");
                 }
             }
-            int posOfWhere = caseString.indexOf("when");
-            String caseForReturn = "case " + caseString.toString().substring(posOfWhere);
-            return new CypReturn(clause.get(1), clause.get(3), COUNT_FALSE, AGG_NONE, caseForReturn, matchC);
+            return new CypReturn(clause.get(1), clause.get(3), COUNT_FALSE, AGG_NONE, caseString.toString(), matchC);
         }
         // 10a.
         else if (cypWalker.hasAverage()) {

### 2017-08-15 src/main/java/intermediate_rep/CypAggFuncs.java
Javadoc update mostly.

@@ -1,15 +1,30 @@
 package intermediate_rep;
 
+/**
+ * Class for mapping the different types of aggregate functions that are possible to return.
+ */
 public class CypAggFuncs {
+    // no aggregate function present in the return item.
     public static int AGG_NONE = 0;
+    // collect() function used.
     public static int AGG_COLLECT = 1;
+    // sum() function used.
     public static int AGG_SUM = 2;
+    // avg() function used.
     public static int AGG_AVG = 3;
+    // min() function used.
     public static int AGG_MIN = 4;
+    // max() function used.
     public static int AGG_MAX = 5;
 
-    public static String convert(int countType) {
-        switch (countType) {
+    /**
+     * Returns a String representation of the type of aggregate function that is present.
+     *
+     * @param aggFuncType The integer stored in the CypReturn object (view CypherTranslator for more information).
+     * @return String meaning of the integer input (usually from the CypReturn getAggFunc() method.)
+     */
+    public static String convert(int aggFuncType) {
+        switch (aggFuncType) {
             case 1:
                 return "collect";
             case 2:
@@ -25,8 +40,16 @@ public class CypAggFuncs {
         }
     }
 
-    public static String sqlEquiv(int countType) {
-        switch (countType) {
+    /**
+     * Returns the SQL equivalent syntax for the aggregate function being used. Mostly this is a straightforward
+     * of the same keyword (i.e. sum() in Cypher unsurprisingly maps to sum() in SQL). collect() in Cypher however
+     * becomes array_agg() in Postgres at least.
+     *
+     * @param aggFuncType The integer stored in the CypReturn object (view CypherTranslator for more information).
+     * @return SQL syntax for the aggregate function represented by the integer argument.
+     */
+    public static String sqlEquiv(int aggFuncType) {
+        switch (aggFuncType) {
             case 1:
                 return "array_agg(";
             case 2:

### 2017-08-15 src/main/java/intermediate_rep/CypCount.java

@@ -1,7 +1,8 @@
 package intermediate_rep;
 
 /**
- * Class for mapping the different types of count() that are possible to return to integers in Java.
+ * Class for mapping the different types of count() that are possible to return. They are represented internally
+ * as integers in Java. Tbe convert() method can be used to get the String interpretation of the integer.
  */
 public class CypCount {
     // no count is present in the return item stored in the CypReturn object.
@@ -13,6 +14,13 @@ public class CypCount {
     // the count function is present, and the field being returned must be distinct.
     public static int COUNT_DISTINCT = 2;
 
+    /**
+     * Returns a String representation of the type of count that is present. If 0 is passed to this function,
+     * then false will be returned. 1 maps to the meaning count(...). 2 maps to the meaning count(distinct ...)
+     *
+     * @param countType The integer stored in the CypReturn object (view CypherTranslator for more information).
+     * @return String meaning of the integer input (usually from the CypReturn getCount() method.)
+     */
     public static String convert(int countType) {
         switch (countType) {
             case 0:

### 2017-08-15 src/main/java/intermediate_rep/CypIterate.java

@@ -1,20 +1,26 @@
 package intermediate_rep;
 
 /**
- * Class for holding intermediate information about the Cypher extension.
- * -- Part II Project specific code. --
+ * Class for holding intermediate information about the Cypher extension that uses the ITERATE keyword.
  */
 
 public class CypIterate {
+    // String equivalent of Cypher input containing the ITERATE keyword.
     private String sql;
+    // the original Cypher input containing the ITERATE keyword.
     private String originalCypherInput;
+    // the initial MATCH part of the input.
     private String firstQuery;
+    // the loop statement of the input.
     private String loopQuery;
+    // the id between the keywords LOOP and ON.
     private String loopIndexTo;
+    // the id between the keywords ON and COLLECT.
     private String loopIndexFrom;
+    // the id after the keyword COLLECT.
     private String collectIndex;
+    // the RETURN statement of the input.
     private String returnStatement;
-
 
     public CypIterate(String cypher) {
         this.originalCypherInput = cypher;

### 2017-08-15 src/main/java/intermediate_rep/CypOrder.java

@@ -26,10 +26,6 @@ public class CypOrder {
         this.ascOrDesc = ascOrDesc;
     }
 
-    public String getNodeId() {
-        return nodeID;
-    }
-
     public String getField() {
         return field;
     }

### 2017-08-15 src/main/java/intermediate_rep/CypWhere.java

@@ -1,8 +1,15 @@
 package intermediate_rep;
 
+/**
+ * Stores information linking to an individual predicate contained within a WHERE clause. For example, information
+ * about a statement a.name = 'Ben' would be stored in one CypWhere object.
+ */
 public class CypWhere {
+    // position in the original Cypher WHERE clause of the component.
     private int posInWhere;
+    // stores string of brackets either before or after component.
     private String bracketing;
+    // stores boolean operator that immediately follows this component (will be either and, or, NULL).
     private String boolOp;
 
     public CypWhere(int id) {

### 2017-08-15 src/main/java/intermediate_rep/DecodedQuery.java

@@ -19,15 +19,15 @@ public class DecodedQuery {
 
     /**
      * Create the DecodedQuery object based on the parsing of the Cypher input.
-     *  @param m     MatchClause object.
+     *
+     * @param m     MatchClause object.
      * @param r     ReturnClause object.
      * @param o     OrderClause object.
      * @param skip  SKIP value.
      * @param limit LIMIT value.
      * @param c     The CypherWalker object.
      */
-    public DecodedQuery(MatchClause m, ReturnClause r, OrderClause o,
-                        int skip, int limit, CypherWalker c) {
+    public DecodedQuery(MatchClause m, ReturnClause r, OrderClause o, int skip, int limit, CypherWalker c) {
         this.matchC = m;
         this.returnC = r;
         this.orderC = o;

### 2017-08-15 src/main/java/intermediate_rep/MatchClause.java

@@ -55,7 +55,7 @@ public class MatchClause {
 
 
     /**
-     * Prints out information about the nodes and relationships in the match clause.
+     * Prints out information about the nodes and relationships in the MATCH clause.
      */
     @Override
     public String toString() {

### 2017-08-15 src/main/java/intermediate_rep/OrderClause.java

@@ -3,15 +3,18 @@ package intermediate_rep;
 import java.util.List;
 
 /**
- * Stores the order clause.
+ * Stores the intermediate representation of the original Cypher ORDER BY clause.
  */
 public class OrderClause {
+    // the individual components of the ORDER BY clause.
     private List<CypOrder> items;
 
+    // returns a list of the individual components.
     public List<CypOrder> getItems() {
         return items;
     }
 
+    // sets the individual components of the ORDER BY clause.
     public void setItems(List<CypOrder> items) {
         this.items = items;
     }

### 2017-08-15 src/main/java/intermediate_rep/ReturnClause.java

@@ -3,21 +3,24 @@ package intermediate_rep;
 import java.util.List;
 
 /**
- * Stores the return clause.
+ * Stores the intermediate representation of the original Cypher RETURN clause.
  */
 public class ReturnClause {
+    // the individual components of the RETURN clause.
     private List<CypReturn> items;
 
+    // returns a list of the individual components.
     public List<CypReturn> getItems() {
         return items;
     }
 
+    // sets the individual components of the RETURN clause.
     public void setItems(List<CypReturn> items) {
         this.items = items;
     }
 
     /**
-     * Prints out information about the individual items in the match clause.
+     * Prints out information about the individual items in the RETURN clause.
      */
     @Override
     public String toString() {

### 2017-08-16 src/main/java/database/Neo4jDriver.java
More sensible exceptions added and more refactoring of code.

@@ -26,6 +26,7 @@ public class Neo4jDriver {
      *
      * @param query          Cypher to execute.
      * @param cypher_results File to store the results.
+     * @param props          C2SProperties object (should already be initialised).
      * @param printOutput    Set to true to store the outputs of the query on disk.
      */
     public static void run(String query, File cypher_results, boolean printOutput, C2SProperties props) {

### 2017-08-16 src/main/java/database/key_value_hazelcast/KeyValueTest.java

@@ -15,8 +15,8 @@ import java.util.*;
 
 /**
  * Main class for inserting a graph schema into a key-value store.
- * This class is mainly used as an example - the key-value store being
- * used is Hazelcast, which is an in-memory database.
+ * This class is mainly used as an example - the key-value store being used is Hazelcast,
+ * which is an in-memory database.
  */
 public class KeyValueTest {
     private static Map<Integer, String> mapNodes;

### 2017-08-16 src/main/java/database/postgres/InsertSchemaPostgres.java

@@ -12,7 +12,7 @@ import java.util.ArrayList;
 import java.util.List;
 
 /**
- * Class containing all methods required to store the results of the schema conversion from Neo4j to Postgres.
+ * Store the results of the schema conversion from Neo4j to Postgres.
  */
 public class InsertSchemaPostgres {
     private static List<String> fieldsForMetaFile = new ArrayList<>();
@@ -21,6 +21,7 @@ public class InsertSchemaPostgres {
      * Executing the various schema parts one by one to Postgres.
      *
      * @param database Name of the Postgres database to store the new schema on.
+     * @param props    C2SProperties object (should already be initialised).
      */
     public static void executeSchemaChange(String database, C2SProperties props) {
         PostgresDriver.createConnection(database, props);
@@ -54,6 +55,8 @@ public class InsertSchemaPostgres {
      * All of the fields and relationships gathered during the schema conversion are stored in
      * meta files (to be used when outputting the results of the queries from both Postgres and
      * Neo4j).
+     *
+     * @param props C2SProperties object (should already be initialised).
      */
     private static void addFieldsToMetaFile(C2SProperties props) {
         FileOutputStream fos;
@@ -86,6 +89,7 @@ public class InsertSchemaPostgres {
      * If there is a label which is applied to a node only ever on its own in isolation, then store this as a
      * relation to remove unnecessary NULLs which slow execution of SQL down.
      *
+     * @param props C2SProperties object (should already be initialised).
      * @return SQL to execute.
      */
     private static String insertEachLabel(C2SProperties props) {
@@ -342,6 +346,16 @@ public class InsertSchemaPostgres {
         return sb;
     }
 
+    /**
+     * For each type of relationship in the original Neo4j graph, add a relation into Postgres.
+     *
+     * @param sbTypes The StringBuilder object containing all of the SQL to be executed.
+     * @param columns The columns of the relation to be committed to the database.
+     * @param o       The JSONObject containing the value of the key 'type'. The value will be the name of the
+     *                relation (for example, if the value is 'OWNS', the relation stored will be 'e$OWNS'.
+     * @param values  String of the values to store into the database.
+     * @return Updated StringBuilder object with additional SQL to execute on the database.
+     */
     private static StringBuilder addType(StringBuilder sbTypes, String columns, JsonObject o, String values) {
         sbTypes.append("INSERT INTO ");
         sbTypes.append("e$").append(o.get("type").getAsString()).append("(").append(columns);
[…]

### 2017-08-16 src/main/java/database/postgres/PostgresConstants.java

@@ -1,7 +1,7 @@
 package database.postgres;
 
 /**
- * RelDBConstants holds strings that are inserted into a relational database during the initial schema conversion.
+ * PostgresConstants holds strings that are inserted into a relational database during the initial schema conversion.
  */
 class PostgresConstants {
     static final String ADJLIST_FROM = "CREATE MATERIALIZED VIEW adjList_from AS (select idl as LeftNode, " +

### 2017-08-16 src/main/java/database/postgres/PostgresDriver.java

@@ -25,6 +25,7 @@ public class PostgresDriver {
      * Create the initial connection to the database.
      *
      * @param dbName Name of the database to connect to.
+     * @param props  C2SProperties object (should already be initialised).
      */
     static void createConnection(String dbName, C2SProperties props) {
         try {
@@ -59,6 +60,7 @@ public class PostgresDriver {
      *
      * @param query  SQL to execute (beginning with CREATE)
      * @param dbName Database name to execute on.
+     * @param props  C2SProperties object (should already be initialised).
      * @throws SQLException Error with the SQL query being executed.
      */
     public static void executeCreateView(String query, String dbName, C2SProperties props) throws SQLException {
@@ -81,6 +83,7 @@ public class PostgresDriver {
      * @param database    Database to execute statement on.
      * @param pg_results  File to store the results.
      * @param printOutput Set to true if the output of the SQL statement should be stored in a local file.
+     * @param props       C2SProperties object (should already be initialised).
      * @throws SQLException Thrown if there is an error in the SQL statement.
      */
     public static void select(String query, String database, File pg_results, boolean printOutput, C2SProperties props)
@@ -190,6 +193,7 @@ public class PostgresDriver {
      *
      * @param query  SQL statement to execute.
      * @param dbName Database name of the database to execute the statement on.
+     * @param props  C2SProperties object (should already be initialised).
      * @throws SQLException Error with the transaction.
      */
     public static void insertOrDelete(String query, String dbName, C2SProperties props) throws SQLException {

### 2017-08-16 src/main/java/exceptions/ConversionSQLException.java

@@ -1,0 +1,7 @@
+package exceptions;
+
+public class ConversionSQLException extends Exception {
+    public ConversionSQLException(String s) {
+        super(s);
+    }
+}

### 2017-08-16 src/main/java/exceptions/DQInvalidException.java

@@ -1,0 +1,7 @@
+package exceptions;
+
+public class DQInvalidException extends Exception {
+    public DQInvalidException(String s) {
+        super(s);
+    }
+}

### 2017-08-16 src/main/java/intermediate_rep/CypReturn.java

@@ -1,5 +1,6 @@
 package intermediate_rep;
 
+import exceptions.DQInvalidException;
 import query_translation.sql.utilities_sql.WithSQL;
 
 /**
@@ -33,10 +34,10 @@ public class CypReturn {
      * @param matchC     The MatchClause object associated with the same Cypher query is also needed, so that
      *                   the type of the return item can be calculated (type meaning either the return item is returning
      *                   a node, a relationship, or is part of a longer Cypher query that contains the WITH keyword.)
-     * @throws Exception Error in discovering the type of the return item.
+     * @throws DQInvalidException Error in discovering the type of the return item.
      */
     public CypReturn(String id, String field, int count_x, int agg_func_x, String caseS, MatchClause matchC)
-            throws Exception {
+            throws DQInvalidException {
         this.nodeID = id;
         this.field = field;
         this.count = count_x;
@@ -69,10 +70,9 @@ public class CypReturn {
      * @param idReturnItem The id of the return item for which the type is being calculated for.
      * @param matchC       The MatchClause object for the Cypher input.
      * @return String - one of ['node', 'rel', 'withNode']
-     * @throws Exception The return item is not a node, relationship, or part of a longer Cypher input, and therefore
-     *                   must be invalid.
+     * @throws DQInvalidException Error in discovering the type of the return item.
      */
-    private String discoverType(String idReturnItem, MatchClause matchC) throws Exception {
+    private String discoverType(String idReturnItem, MatchClause matchC) throws DQInvalidException {
         //check the nodes first
         for (CypNode cN : matchC.getNodes()) {
             if (cN.getId() != null && cN.getId().equals(idReturnItem)) {
@@ -97,7 +97,7 @@ public class CypReturn {
             }
         }
 
-        throw new Exception("RETURN DISCOVER TYPE INCORRECT - NOT A NODE OR REL");
+        throw new DQInvalidException("Failed to discover the return type - not a node or a relationship...");
     }
 
     public String getNodeID() {

### 2017-08-16 src/main/java/intermediate_rep/DecodedQuery.java

@@ -2,6 +2,8 @@ package intermediate_rep;
 
 import translator.CypherWalker;
 
+import java.util.ArrayList;
+
 /**
  * This class definition bundles together all of the intermediate representations
  * built up by the tool, as well as storing the SQL translation for completeness.
@@ -11,11 +13,14 @@ public class DecodedQuery {
     private ReturnClause returnC;
     private OrderClause orderC;
     private CypForEach forEachC;
+    private CypIterate iterate;
     private int skipAmount;
     private int limitAmount;
     private CypherWalker cypherAdditionalInfo;
     // The SQL translation is also stored in this object for completeness.
     private String sqlEquiv;
+    private ArrayList<String> withParts = new ArrayList<>();
+    private ArrayList<DecodedQuery> unionParts = new ArrayList<>();
 
     /**
      * Create the DecodedQuery object based on the parsing of the Cypher input.
@@ -27,13 +32,19 @@ public class DecodedQuery {
      * @param limit LIMIT value.
      * @param c     The CypherWalker object.
      */
-    public DecodedQuery(MatchClause m, ReturnClause r, OrderClause o, int skip, int limit, CypherWalker c) {
+    public DecodedQuery(MatchClause m, ReturnClause r, OrderClause o, int skip, int limit, CypherWalker c,
+                        CypIterate it) {
         this.matchC = m;
         this.returnC = r;
         this.orderC = o;
         this.skipAmount = skip;
         this.limitAmount = limit;
         this.cypherAdditionalInfo = c;
+        this.iterate = it;
+    }
+
+    public DecodedQuery() {
+
     }
 
     public MatchClause getMc() {
@@ -74,5 +85,25 @@ public class DecodedQuery {
 
     public void setForEachC(CypForEach forEachC) {
         this.forEachC = forEachC;
+    }
+
+    public CypIterate getIterate() {
+        return iterate;
+    }
+
+    public ArrayList<String> getWithParts() {
+        return withParts;
+    }
+
+    public void setWithParts(ArrayList<String> withParts) {
+        this.withParts = withParts;
+    }
+
+    public void addToUnionParts(DecodedQuery unionDQ) {
+        unionParts.add(unionDQ);
+    }
+
+    public ArrayList<DecodedQuery> getUnionParts() {
+        return unionParts;
     }
 }

### 2017-08-16 src/main/java/production/C2SInteractive.java

@@ -14,11 +14,12 @@ class C2SInteractive {
      *
      * @param f_cypher File object to store the results from Neo4j.
      * @param f_sql    File object to store the results from Postgres.
+     * @param props    C2SProperties object (should already be initialised).
      * @param dbName   Name of the relational database that the generated SQL will be executed on.
      */
     static void run_debug(File f_cypher, File f_sql, String dbName, C2SProperties props) {
         System.out.println("PRINT TO FILE : " + ((C2SMain.printBool) ? "enabled" : "disabled"));
-        System.out.println("Cypher to SQL Translator Tool v1.0");
+        System.out.println("Cypher to SQL Translator Tool v1.1");
         System.out.println("To exit, type :exit.");
 
         while (true) {
@@ -33,7 +34,6 @@ class C2SInteractive {
                 try {
                     C2SMain.translateCypherToSQL(resp, f_cypher, f_sql, dbName, true, props);
                 } catch (Exception e) {
-                    System.err.println("Error with the last query...");
                     e.printStackTrace();
                 }
             }
@@ -45,9 +45,10 @@ class C2SInteractive {
      * the results are outputted to the console, but from the relational database.
      *
      * @param dbName Name of the relational database that the translated query is to be executed on.
+     * @param props  C2SProperties object (should already be initialised).
      */
     static void run(String dbName, C2SProperties props) {
-        System.out.println("Cypher to SQL Translator Tool v1.0");
+        System.out.println("Cypher to SQL Translator Tool v1.1");
         System.out.println("To exit, type :exit.");
 
         while (true) {
@@ -62,7 +63,6 @@ class C2SInteractive {
                 try {
                     C2SMain.translateCypherToSQL(resp, null, null, dbName, false, props);
                 } catch (Exception e) {
-                    System.err.println("Error with the last query...");
                     e.printStackTrace();
                 }
             }

### 2017-08-16 src/main/java/production/C2SMain.java

@@ -3,6 +3,8 @@ package production;
 import database.Neo4jDriver;
 import database.postgres.InsertSchemaPostgres;
 import database.postgres.PostgresDriver;
+import exceptions.ConversionSQLException;
+import exceptions.DQInvalidException;
 import intermediate_rep.DecodedQuery;
 import org.apache.commons.io.FileUtils;
 import org.neo4j.driver.v1.exceptions.ClientException;
@@ -17,20 +19,18 @@ import java.util.HashMap;
 import java.util.Map;
 
 /**
- * Main class for starting the application.
- * View README.md for additional guidance.
+ * Main class for starting the application. View README.md for usage.
  */
 public class C2SMain {
     // for optimisations based on the return clause of Cypher
     public static final Map<String, String> labelProps = new HashMap<>();
     // for use in deleting relationships attached to nodes
     public static final ArrayList<String> allRelTypes = new ArrayList<>();
+
     private static final String OS = System.getProperty("os.name").toLowerCase();
     public static int numResultsNeo4j;
     public static int numResultsPostgres;
-    // stores the current DecodedQuery object so that the Cypher results module can use it,
-    // without having to rerun computation.
-    public static DecodedQuery currentDQ = null;
+
     // static variable set by the translation unit if the ID needs to be returned from the Postgres database
     // by default it is set to false as this is the same logic as Neo4j
     public static boolean needToPrintID = false;
@@ -108,7 +108,7 @@ public class C2SMain {
      * - Secondly, it reads in all the possible types of relationships
      * (used for example when deleting nodes from the relational schema).
      *
-     * @param props
+     * @param props C2SProperties object.
      */
     static void getLabelMapping(C2SProperties props) {
         // open file and read in property keys, removing duplicates as they are of no use.
@@ -159,6 +159,8 @@ public class C2SMain {
      * https://neo4j.com/developer/kb/warm-the-cache-to-improve-performance-from-cold-start/
      * also, if there is an SSL issue when changing databases, attempt to fix it by running
      * some fix up code.
+     *
+     * @param props C2SProperties object.
      */
     static void warmUpResetSSL(C2SProperties props) {
         try {
@@ -215,6 +217,7 @@ public class C2SMain {
      * on a relational backend.
      *
      * @param databaseName The name of the database to be used for storing the relational representation.
+     * @param props        C2SProperties object.
      */
[…]

### 2017-08-16 src/main/java/query_translation/sql/conversion_types/AbstractConversion.java

@@ -1,12 +1,13 @@
 package query_translation.sql.conversion_types;
 
+import exceptions.DQInvalidException;
 import intermediate_rep.DecodedQuery;
-import production.C2SMain;
 import production.C2SProperties;
 import query_translation.sql.utilities_sql.SQLTranslate;
 import query_translation.sql.utilities_sql.UnionSQL;
 import translator.CypherTokenizer;
 
+import java.io.IOException;
 import java.util.ArrayList;
 
 /**
@@ -19,57 +20,95 @@ public abstract class AbstractConversion {
      * Convert Cypher queries to SQL
      *
      * @param cypher Original Cypher query to translate.
+     * @param props  properties object.
      * @return DecodedQuery object containing all the necessary intermediate representation of the Cypher input.
+     * @throws DQInvalidException The Cypher could not be translated to a DecodedQuery object.
      */
-    public static DecodedQuery convertCypherToSQL(String cypher, C2SProperties props) {
+    public static DecodedQuery genDQAndSQL(String cypher, C2SProperties props) throws DQInvalidException {
         // A DecodedQuery object stores the intermediate representation of the Cypher input.
         // In theory, this same DecodedQuery object could then be translated to different backend implementations.
-        DecodedQuery dQ = null;
+        DecodedQuery dQ = new DecodedQuery();
 
-        try {
-            if (cypher.toLowerCase().contains(" union all ")) {
-                String[] queries = cypher.toLowerCase().split(" union all ");
-                ArrayList<String> unionSQL = new ArrayList<>();
+        if (cypher.toLowerCase().contains(" union all ")) {
+            String[] queries = cypher.toLowerCase().split(" union all ");
+            ArrayList<String> unionSQL = new ArrayList<>();
 
-                for (String s : queries) {
-                    dQ = CypherTokenizer.decode(s, false);
-                    unionSQL.add(SQLTranslate.translateRead(dQ, props));
-                }
-
-                dQ.setSqlEquiv(UnionSQL.genUnion(unionSQL, "UNION ALL"));
-            } else if (cypher.toLowerCase().contains(" union ")) {
-                String[] queries = cypher.toLowerCase().split(" union ");
-                ArrayList<String> unionSQL = new ArrayList<>();
-
-                for (String s : queries) {
-                    dQ = CypherTokenizer.decode(s, false);
-                    unionSQL.add(SQLTranslate.translateRead(dQ, props));
-                }
-
[…]

### 2017-08-16 src/main/java/query_translation/sql/conversion_types/ForEach_Cypher.java

@@ -1,5 +1,6 @@
 package query_translation.sql.conversion_types;
 
+import exceptions.DQInvalidException;
 import intermediate_rep.CypForEach;
 import intermediate_rep.DecodedQuery;
 import production.C2SProperties;
@@ -12,19 +13,23 @@ import query_translation.sql.utilities_sql.ForEach;
  */
 public class ForEach_Cypher extends AbstractConversion {
     @Override
-    public String convertQuery(String cypher, C2SProperties props) {
+    public String convertToSQL(DecodedQuery dQ, C2SProperties props) {
+        ForEach fe = new ForEach();
+        return fe.translate(new StringBuilder(), dQ, props).toString();
+    }
+
+    @Override
+    public DecodedQuery generateDQ(String cypher, C2SProperties props) throws DQInvalidException {
         String changeLine = cypher.toLowerCase().replace("with", "return");
         String[] feParts = changeLine.toLowerCase().split(" foreach ");
 
-        DecodedQuery dQ = convertCypherToSQL(feParts[0].trim() + ";", props);
+        DecodedQuery dQ = genDQAndSQL(feParts[0].trim() + ";", props);
 
         CypForEach cypForEach = new CypForEach(feParts[1].trim());
         if (dQ != null) {
             dQ.setForEachC(cypForEach);
-        }
+        } else throw new DQInvalidException("Cypher containing FOREACH keyword is invalid.");
 
-        StringBuilder sql = new StringBuilder();
-        ForEach fe = new ForEach();
-        return fe.translate(sql, dQ, props).toString();
+        return dQ;
     }
 }

### 2017-08-16 src/main/java/query_translation/sql/conversion_types/Iterate_Cypher.java

@@ -1,6 +1,8 @@
 package query_translation.sql.conversion_types;
 
+import exceptions.DQInvalidException;
 import intermediate_rep.CypIterate;
+import intermediate_rep.DecodedQuery;
 import production.C2SProperties;
 import query_translation.sql.utilities_sql.IterateSQL;
 
@@ -11,7 +13,13 @@ import query_translation.sql.utilities_sql.IterateSQL;
  */
 public class Iterate_Cypher extends AbstractConversion {
     @Override
-    public String convertQuery(String cypher, C2SProperties props) {
+    public String convertToSQL(DecodedQuery dQ, C2SProperties props) throws DQInvalidException {
+        IterateSQL it = new IterateSQL();
+        return it.translate(new StringBuilder(), dQ, props).toString();
+    }
+
+    @Override
+    public DecodedQuery generateDQ(String cypher, C2SProperties props) {
         CypIterate cypIter = new CypIterate(cypher);
 
         String line = cypIter.getOriginalCypherInput();
@@ -29,8 +37,6 @@ public class Iterate_Cypher extends AbstractConversion {
         String firstStep = matchClause + "return " + cypIter.getLoopIndexTo() + ";";
         cypIter.setFirstQuery(firstStep);
 
-        StringBuilder sql = new StringBuilder();
-        IterateSQL it = new IterateSQL();
-        return it.translate(sql, cypIter, props);
+        return new DecodedQuery(null, null, null, -1, -1, null, cypIter);
     }
 }

### 2017-08-16 src/main/java/query_translation/sql/conversion_types/Multiple_With_Cypher.java

@@ -1,5 +1,6 @@
 package query_translation.sql.conversion_types;
 
+import exceptions.DQInvalidException;
 import intermediate_rep.DecodedQuery;
 import production.C2SProperties;
 import query_translation.sql.utilities_sql.WithSQL;
@@ -15,31 +16,21 @@ public class Multiple_With_Cypher extends AbstractConversion {
     public static Map<String, String> mappingMultipleWith;
 
     @Override
-    public String convertQuery(String cypher, C2SProperties props) {
+    public String convertToSQL(DecodedQuery dQ, C2SProperties props) throws DQInvalidException {
         mappingMultipleWith = new HashMap<>();
 
-        ArrayList<String> withParts = new ArrayList<>();
-        String cypherCopy = cypher.toLowerCase();
-
-        while (!cypherCopy.isEmpty()) {
-            int lastMatchIndex = cypherCopy.lastIndexOf("match");
-            withParts.add(cypherCopy.substring(lastMatchIndex, cypherCopy.length() - 1) + ";");
-            cypherCopy = cypherCopy.substring(0, lastMatchIndex);
-        }
-
-        int numParts = withParts.size();
+        int numParts = dQ.getWithParts().size();
         StringBuilder multipleWithSQL = new StringBuilder();
 
         for (int i = 0; i < numParts - 1; i++) {
-            String withQuery = withParts.get(numParts - (i + 1)).replace(" with ", " return ");
-            DecodedQuery withPartDQ = convertCypherToSQL(withQuery, props);
+            String withQuery = dQ.getWithParts().get(numParts - (i + 1)).replace(" with ", " return ");
+            DecodedQuery withPartDQ = genDQAndSQL(withQuery, props);
             mappingMultipleWith.put(withPartDQ.getRc().getItems().get(0).getNodeID(),
                     "w" + String.valueOf(alphabet[i]).toUpperCase());
             multipleWithSQL.append(WithSQL.genTemp(withPartDQ.getSqlEquiv(), i, withPartDQ)).append(" ");
         }
 
-        // System.out.println(multipleWithSQL);
-        DecodedQuery finalPartDQ = convertCypherToSQL(withParts.get(0), props);
+        DecodedQuery finalPartDQ = genDQAndSQL(dQ.getWithParts().get(0), props);
         String finalSQL = finalPartDQ.getSqlEquiv();
 
         // if ORDER BY and/or LIMIT/SKIP have been appended, temporarily remove them, then add back after
[…]

### 2017-08-16 src/main/java/query_translation/sql/conversion_types/SP_Cypher.java

@@ -1,7 +1,7 @@
 package query_translation.sql.conversion_types;
 
+import exceptions.DQInvalidException;
 import intermediate_rep.DecodedQuery;
-import production.C2SMain;
 import production.C2SProperties;
 import query_translation.sql.utilities_sql.ShortestPath;
 import translator.CypherTokenizer;
@@ -11,10 +11,19 @@ import translator.CypherTokenizer;
  */
 public class SP_Cypher extends AbstractConversion {
     @Override
-    public String convertQuery(String cypher, C2SProperties props) {
+    public String convertToSQL(DecodedQuery dQ, C2SProperties props) {
+        ShortestPath sp = new ShortestPath();
+        return sp.translate(new StringBuilder(), dQ, props).toString();
+    }
+
+    @Override
+    public DecodedQuery generateDQ(String cypher, C2SProperties props) throws DQInvalidException {
         cypher = cypher.toLowerCase();
         int returnIndex = cypher.indexOf("return");
+        if (returnIndex == -1)
+            throw new DQInvalidException("Cypher input has no RETURN clause --> malformed.");
         int whereIndex = cypher.indexOf("where");
+
         String whereClause = null;
 
         if (whereIndex != -1) {
@@ -27,17 +36,9 @@ public class SP_Cypher extends AbstractConversion {
         String returnClause = cypher.substring(cypher.indexOf("return"));
         String cypherPathQuery = "MATCH " + path + ((whereIndex != -1) ? whereClause : "") + " " + returnClause;
 
-        DecodedQuery dQMainPath;
-        try {
-            dQMainPath = CypherTokenizer.decode(cypherPathQuery, false);
-        } catch (Exception e) {
-            e.printStackTrace();
-            return null;
-        }
+        DecodedQuery dQ = CypherTokenizer.decode(cypherPathQuery, false);
 
-        C2SMain.currentDQ = dQMainPath;
-        StringBuilder sql = new StringBuilder();
-        ShortestPath sp = new ShortestPath();
-        return sp.translate(sql, dQMainPath, props).toString();
+        if (dQ == null) throw new DQInvalidException("Could not convert Cypher input...");
+        return dQ;
     }
 }

### 2017-08-16 src/main/java/query_translation/sql/conversion_types/With_Cypher.java

@@ -1,5 +1,6 @@
 package query_translation.sql.conversion_types;
 
+import exceptions.DQInvalidException;
 import intermediate_rep.DecodedQuery;
 import production.C2SProperties;
 import query_translation.sql.utilities_sql.WithSQL;
@@ -14,16 +15,21 @@ public class With_Cypher extends AbstractConversion {
      * @param cypher Original Cypher input containing the WITH keyword.
      * @return SQL string equivalent of the original Cypher input.
      */
-    @Override
     public String convertQuery(String cypher, C2SProperties props) {
         int posOfMatch = cypher.toLowerCase().lastIndexOf("match");
         int posOfWhere = cypher.toLowerCase().indexOf("where");
         int posOfOrderBy = cypher.toLowerCase().indexOf("order by");
 
         // from the tokens in the original Cypher query, decide on the most appropriate method for translation.
-        if (posOfMatch != -1 && posOfMatch != 0) return withMatch(cypher, props);
-        if (posOfOrderBy == -1 || (posOfWhere != -1 && (posOfWhere < posOfOrderBy))) return withWhere(cypher, props);
-        else return withOB(cypher, props);
+        try {
+            if (posOfMatch != -1 && posOfMatch != 0) return withMatch(cypher, props);
+            if (posOfOrderBy == -1 || (posOfWhere != -1 && (posOfWhere < posOfOrderBy)))
+                return withWhere(cypher, props);
+            else return withOB(cypher, props);
+        } catch (DQInvalidException ex) {
+            ex.printStackTrace();
+            return null;
+        }
     }
 
     /**
@@ -34,11 +40,11 @@ public class With_Cypher extends AbstractConversion {
      * @param props
      * @return SQL equivalent of the Cypher input.
      */
-    private String withMatch(String cypher, C2SProperties props) {
+    private String withMatch(String cypher, C2SProperties props) throws DQInvalidException {
         String changeLine = cypher.toLowerCase().replace("with", "return");
         String firstWith = changeLine.toLowerCase()
                 .substring(0, cypher.toLowerCase().lastIndexOf("match") + 1) + ";";
-        DecodedQuery dqFirstWith = convertCypherToSQL(firstWith, props);
+        DecodedQuery dqFirstWith = genDQAndSQL(firstWith, props);
 
         String withTemp = null;
         if (dqFirstWith != null) {
[…]

### 2017-08-16 src/main/java/query_translation/sql/utilities_sql/AbstractTranslation.java

@@ -1,5 +1,6 @@
 package query_translation.sql.utilities_sql;
 
+import exceptions.DQInvalidException;
 import intermediate_rep.DecodedQuery;
 import production.C2SProperties;
 
@@ -13,5 +14,5 @@ abstract class AbstractTranslation {
     static char[] extendID = "123456789".toCharArray();
     static boolean usesOptimalTable = false;
 
-    public abstract StringBuilder translate(StringBuilder sql, DecodedQuery decodedQuery, C2SProperties props);
+    public abstract StringBuilder translate(StringBuilder sql, DecodedQuery decodedQuery, C2SProperties props) throws DQInvalidException;
 }

### 2017-08-16 src/main/java/query_translation/sql/utilities_sql/IterateSQL.java

@@ -1,6 +1,7 @@
 package query_translation.sql.utilities_sql;
 
 import com.google.gson.JsonObject;
+import exceptions.DQInvalidException;
 import intermediate_rep.CypIterate;
 import intermediate_rep.CypNode;
 import intermediate_rep.DecodedQuery;
@@ -18,8 +19,10 @@ public class IterateSQL extends AbstractTranslation {
         return -1;
     }
 
-    public String translate(StringBuilder sql, CypIterate ci, C2SProperties props) {
+    @Override
+    public StringBuilder translate(StringBuilder sql, DecodedQuery dQ, C2SProperties props) throws DQInvalidException {
         // get correct loop query
+        CypIterate ci = dQ.getIterate();
         String firstStep = ci.getFirstQuery();
         int posOfLoopFrom = firstStep.indexOf(ci.getLoopIndexFrom());
         String lQuery = "MATCH " + firstStep.substring(posOfLoopFrom - 1, firstStep.length());
@@ -27,10 +30,10 @@ public class IterateSQL extends AbstractTranslation {
 
         // generate the traditional translation to SQL for the loop query (store in string as used
         // multiple times)
-        DecodedQuery loopDQ = AbstractConversion.convertCypherToSQL(ci.getFirstQuery(), props);
+        DecodedQuery loopDQ = AbstractConversion.genDQAndSQL(ci.getFirstQuery(), props);
         String loopSQL = loopDQ.getSqlEquiv();
 
-        String returnSQL = AbstractConversion.convertCypherToSQL(ci.getReturnStatement(), props).getSqlEquiv();
+        String returnSQL = AbstractConversion.genDQAndSQL(ci.getReturnStatement(), props).getSqlEquiv();
         returnSQL = returnSQL.substring(0, returnSQL.length() - 1);
 
         // need to modify loopSQL for the main SQL statement.
@@ -44,7 +47,7 @@ public class IterateSQL extends AbstractTranslation {
         // create the loop_work function with this string
         String loopWorkStr;
 
-        loopDQ = AbstractConversion.convertCypherToSQL(ci.getLoopQuery(), props);
+        loopDQ = AbstractConversion.genDQAndSQL(ci.getLoopQuery(), props);
         int posLoopFrom = calculatePos(ci.getLoopIndexFrom(), loopDQ);
         JsonObject obj = loopDQ.getMc().getNodes().get(posLoopFrom - 1).getProps();
         if (obj == null) {
[…]

### 2017-08-16 src/main/java/query_translation/sql/utilities_sql/SQLTranslate.java

@@ -1,5 +1,6 @@
 package query_translation.sql.utilities_sql;
 
+import exceptions.DQInvalidException;
 import intermediate_rep.*;
 import production.C2SMain;
 import production.C2SProperties;
@@ -23,12 +24,13 @@ import static intermediate_rep.CypCount.COUNT_FALSE;
  * Agnostic to the methods above is appending the ORDER BY, GROUP BY, LIMIT and SKIP elements.
  */
 public class SQLTranslate {
-    public static String translateRead(DecodedQuery decodedQuery, C2SProperties props) throws Exception {
+    public static String translateRead(DecodedQuery decodedQuery, C2SProperties props)
+            throws DQInvalidException, IOException {
         // SQL built up from a StringBuilder object.
         StringBuilder sql = new StringBuilder();
 
-        if (decodedQuery.getMc().getNodes().isEmpty()) throw new Exception("MATCH CLAUSE INVALID");
-        if (decodedQuery.getRc().getItems() == null) throw new Exception("RETURN CLAUSE INVALID");
+        if (decodedQuery.getMc().getNodes().isEmpty()) throw new DQInvalidException("MATCH CLAUSE INVALID");
+        if (decodedQuery.getRc().getItems() == null) throw new DQInvalidException("RETURN CLAUSE INVALID");
 
         // There are 3 Cypher query types that can be translated:
         // - queries with no relations in
@@ -237,8 +239,8 @@ public class SQLTranslate {
      * @return Query after GROUP BY
      * @throws IOException Error reading the associated metafile from the workarea location.
      */
-    private static StringBuilder obtainGroupByClause(ReturnClause rc, StringBuilder sql,
-                                                     C2SProperties props) throws IOException {
+    private static StringBuilder obtainGroupByClause(ReturnClause rc, StringBuilder sql, C2SProperties props)
+            throws IOException {
         sql.append(" GROUP BY ");
 
         int nodeTableCount = 0;

### 2017-08-16 src/main/java/query_translation/sql/utilities_sql/WithSQL.java

@@ -1,5 +1,6 @@
 package query_translation.sql.utilities_sql;
 
+import exceptions.DQInvalidException;
 import intermediate_rep.CypAggFuncs;
 import intermediate_rep.CypReturn;
 import intermediate_rep.DecodedQuery;
@@ -184,11 +185,12 @@ public class WithSQL {
         return sWith.toString();
     }
 
-    public static String createSelectMatch(String secondWith, DecodedQuery dqFirstWith, C2SProperties props) {
+    public static String createSelectMatch(String secondWith, DecodedQuery dqFirstWith, C2SProperties props)
+            throws DQInvalidException {
         withMapping.put(dqFirstWith.getRc().getItems().get(0).getNodeID(), "wA");
 
         StringBuilder resSecWith = new StringBuilder();
-        DecodedQuery dQSecWith = AbstractConversion.convertCypherToSQL(secondWith, props);
+        DecodedQuery dQSecWith = AbstractConversion.genDQAndSQL(secondWith, props);
 
         if (dQSecWith.getMc().getRels().isEmpty()) {
             NoRels nr = new NoRels();

### 2017-08-16 src/main/java/translator/CypherTokenizer.java

@@ -25,7 +25,7 @@ public class CypherTokenizer {
      * @return DecodedQuery object representing the Cypher input.
      * @throws Exception Error somewhere.
      */
-    public static DecodedQuery decode(String cypher, boolean DEBUG_PRINT) throws Exception {
+    public static DecodedQuery decode(String cypher, boolean DEBUG_PRINT) {
         return CypherTranslator.generateDecodedQuery(getTokenList(cypher, DEBUG_PRINT), cypherWalker);
     }
 

### 2017-08-16 src/main/java/translator/CypherTranslator.java

@@ -2,6 +2,7 @@ package translator;
 
 import com.google.gson.JsonObject;
 import com.google.gson.JsonParser;
+import exceptions.DQInvalidException;
 import intermediate_rep.*;
 import production.C2SMain;
 
@@ -27,9 +28,8 @@ class CypherTranslator {
      * @param tokenList Token list generated from the ANTLR framework.
      * @param cypWalker CypherWalker object containing additional info about the Cypher input.
      * @return DecodedQuery object representing the Cypher input.
-     * @throws Exception Failed to generate a valid DecodedObject for the Cypher input.
      */
-    static DecodedQuery generateDecodedQuery(ArrayList<String> tokenList, CypherWalker cypWalker) throws Exception {
+    static DecodedQuery generateDecodedQuery(ArrayList<String> tokenList, CypherWalker cypWalker) {
         // find positions of the tokens in the query (-1 means not found)
         int posOfMatch = tokenList.indexOf("match");
         int posOfWhere = tokenList.indexOf("where");
@@ -37,7 +37,11 @@ class CypherTranslator {
 
         // if no RETURN keyword is present, then the Cypher input must contain a CREATE OR DELETE keyword.
         if (posOfReturn == -1) {
-            return generateDQCreateDelete(tokenList, cypWalker, posOfMatch, posOfWhere);
+            try {
+                return generateDQCreateDelete(tokenList, cypWalker, posOfMatch, posOfWhere);
+            } catch (DQInvalidException ex) {
+                return null;
+            }
         } else {
             int posOfOrder = tokenList.indexOf("order");
             int posOfSkip = tokenList.indexOf("skip");
@@ -83,21 +87,26 @@ class CypherTranslator {
                 }
             }
 
-            MatchClause matchC = matchDecode(matchClause);
-            ReturnClause returnC = returnDecode(returnClause, matchC, cypWalker);
+            try {
+                MatchClause matchC = matchDecode(matchClause);
+                ReturnClause returnC = returnDecode(returnClause, matchC, cypWalker);
 
-            // if ORDER BY is present in the query
-            OrderClause orderC = null;
-            if (orderClause != null) {
-                orderC = orderDecode(orderClause);
+                // if ORDER BY is present in the query
+                OrderClause orderC = null;
+                if (orderClause != null) {
+                    orderC = orderDecode(orderClause);
+                }
+
[…]

### 2017-08-16 docs/latex/draftB.tex
Improved and more accurate LaTeX documentation added.

@@
 {
-backgroundcolor=\color{black},
+backgroundcolor=\color{gray},
 basicstyle=\color{white}\ttfamily,
@@
 
+\lstset{frame=tb,
+language=Java,
+aboveskip=3mm,
+belowskip=3mm,
+showstringspaces=false,
+columns=flexible,
+basicstyle={\small\ttfamily},
+numbers=none,
+numberstyle=\tiny\color{gray},
+keywordstyle=\color{blue},
+commentstyle=\color{dkgreen},
+stringstyle=\color{mauve},
+breaklines=true,
+breakatwhitespace=true,
+tabsize=3
+}
+
 \begin{document}
@@
 a different database schema, and different database language (the original aim of this project was to translate
-Cypher to SQL).}
+Cypher to SQL).
+The tool can either be used as a standalone application, or, by utilising the .jar of the Java application, it can be used as an API for other applications.
+The license of the code is to be confirmed.}
 
@@
 \begin{itemize}
-\item A \emph{Schema Conversion} module for creating the initial database in an engine other than Neo4j (such as a relational schema in Postgres)
+\item A \emph{Schema Conversion} module for creating the initial database in a representation other than Neo4j (such as a relational schema in Postgres)
 \item A \emph{Parsing Cypher} module, which is the first part of the translation to another database language (such as SQL).
 This module relies on the ANTLR framework, and other hand-written parsing code
-\item An \emph{Intermediate Representation} is then built in Java
-\item A \emph{Translation Module} takes this intermediate representation, and converts it to the target database language (the tool was originally designed with SQL in mind)
-\item The \emph{Output Results \& Information} module executes both the original Cypher on Neo4j, and the newly created query on the other chosen database(s).
-The execution times of the queries may be recorded, and the results of the queries may be written to a local file if requested.
-In the latter case, the tool can also verify that both the number of results and the results themselves match (this is useful for debugging).
+\item An \emph{Intermediate Representation} is then built in Java -- the result of this stage of the pipeline is a \texttt{DecodedQuery} object.
+\item A \emph{Translation Module} takes this intermediate representation, and converts it to the target database language
+\item The \emph{Output Results \& Information} module can perform many different tasks, depending on the configuration desired:
+\begin{itemize}
+\item It can execute both the original Cypher on Neo4j, and the newly created query on the target database, and record the execution times of the queries.
+The tool may also be made to write the results to local files, so that the outputs can be compared for accuracy.
+\item The tool can also just execute the translated query on the database engine, and parse the results back into the \emph{Cyp2SQL} tool.
+\end{itemize}
 \end{itemize}
@@
 \subsection{Quick usage}
+Using the \texttt{cyp2sql-next.jar} file:
 \begin{verbatim}
@@
 \begin{tabular}{ p{3cm} p{10.5cm} }
--s \textbar{} -schema & Runs the tool for schema conversion. \\
--t \textbar{} -translate & Runs the tool for query translation. \\
-propertiesFile & Path of the properties file \texttt{c2s\_props.properties}. See \S \ref{ss:prop}. \\
-targetDatabaseName & Name of the relational database for either the converted graph schema to be stored into, or for the translated queries to execute on. \\
-dp \textbar{} -dn \textbar{} -r & \textbf{Only needed when in query translation mode}. See \S \ref{sec:tranQ}.
+\texttt{-s \textbar{} -schema} & Runs the tool for schema conversion. See \S \ref{sec:schemaConv}. \\
+\texttt{-t \textbar{} -translate} & Runs the tool for query translation. See \S \ref{sec:tranQ}. \\
+\texttt{propertiesFile} & Path of the properties file \texttt{c2s\_props.properties}. See \S \ref{ss:prop}. \\
+\texttt{targetDatabaseName} & Name of the relational database for either the converted graph schema to be stored into, or for the translated queries to execute on. \\
+\texttt{dp \textbar{} -dn \textbar{} -r} & \textbf{Only needed when in query translation mode}.
 \end{tabular}
 \end{center}
+
+\medskip
+
+Using the code within another application as a library is also possible. Example usage:
+
+\begin{lstlisting}[language=Java, breaklines=true]
+import database.postgres.InsertSchemaPostgres;
+import intermediate_rep.DecodedQuery;
+import production.C2SMain;
+import production.C2SProperties;
+import schema_conversion.SchemaConvert;
+
+import java.io.IOException;
+
+public class C2STestUsage {
+public static void main(String args[]) {
+// these need to be set before anything can occur.
+String propsFile = "C:/Users/ocraw/IdeaProjects/cyp2sql-next/c2s_props.properties";
+C2SProperties props = new C2SProperties(propsFile);
+
+// name of the blank database to either:
+// - convert the schema too
+// - execute the translated Cypher on
+String dbName = "testb";
+
+String thingToDo = "translate";
+
+switch (thingToDo) {
+case "convert":
+// the properties file must be edited first.
+// a dump from Neo4j is also needed.
+boolean successConvert = SchemaConvert.translate(props);
+if (successConvert) InsertSchemaPostgres.executeSchemaChange(dbName, props);
+break;
+case "translate":
+// location of the script to allow results from Postgres to be
+// piped back to this class. Adapt the scripts if necessary.
+String scriptLoc = "C:/Users/ocraw/IdeaProjects/cyp2sql-next/pgdbPlay.bat";
+
+// Cypher query to translate and then execute.
+String cypher = "MATCH (a:Meta) RETURN DISTINCT a.type " +
+"UNION MATCH (a:Process) RETURN DISTINCT a.type;";
+
+try {
+// obtain the intermediate representation
+DecodedQuery dQ = C2SMain.getDQ(cypher, props);
+System.out.println(dQ.getUnionParts().get(0).getMc());
+
+// convert the intermediate representation to SQL
+String sql = C2SMain.getTranslation(cypher, dQ, props);
+System.out.println(sql);
+
+// execute directly on Postgres if desired (the script will pipe
+// the results back into this tool).
+String postgresOutput = C2SMain.runPostgres(sql, dbName, scriptLoc);
+System.out.println(postgresOutput);
+} catch (IOException e) {
+e.printStackTrace();
+}
+break;
+default:
+System.err.println("Not a valid option...");
+System.exit(1);
+}
+}
+}
+\end{lstlisting}
 
@@
 \section{Schema conversion}
+\label{sec:schemaConv}
 The first step of the pipeline in translating Cypher to SQL is to convert the Neo4j graph database to a relational database.
@@
 
-\textcolor{red}{NOTE: the script to run Postgres on Unix systems may not have the right permissions to run with the tool. \texttt{chmod +x auto\_run\_pg.sh} should fix the issue.}
+\textcolor{red}{NOTE: the script to run Postgres on Unix systems when the tool is in \texttt{-r} mode may not have the right permissions to run with the tool.
+\texttt{chmod +x auto\_run\_pg.sh} should fix the issue.}
 
@@
 \bigskip
-\newpage
 
@@
 
+Figure \ref{iterGraphFig} shows a common pattern, \texttt{BLUE --> GREEN --> RED --> BLUE}.
+Finding all the blue nodes at the end of this pattern \emph{iteratively} would be hard to do with the existing Cypher syntax in a concise and clear manner.
+It is a more tractable problem with the \texttt{ITERATE} syntax:
+
+\begin{lstlisting}[language = Cypher]
+ITERATE MATCH (a:l1)-->(b:l2)-->(c:l3)-->(d:l1)
+LOOP d ON a COLLECT n
+RETURN n.uid;
+\end{lstlisting}
+
 \begin{figure}[h]
@@
 \newpage
-Figure \ref{iterGraphFig} shows a common pattern, \texttt{BLUE --> GREEN --> RED --> BLUE}.
-Finding all the blue nodes at the end of this pattern \emph{iteratively} would be hard to do with the existing Cypher syntax in a concise and clear manner.
-It is a more tractable problem with the \texttt{ITERATE} syntax:
-
-\begin{lstlisting}[language = Cypher]
-ITERATE MATCH (a:l1)-->(b:l2)-->(c:l3)-->(d:l1)
-LOOP d ON a COLLECT n
-RETURN n.uid;
-\end{lstlisting}
 
@@
 
+\newpage
+
 \subsection*{\texttt{MATCH ... WHERE ... RETURN}}
@@
 
+\newpage
+
 \subsection*{\texttt{UNION [ALL]?}}
@@
 
-\newpage
-
 \subsection*{Variable length paths -- \texttt{-[*x..y]->}}
@@
 \end{lstlisting}
+
+\newpage
 

### 2017-08-17 src/main/java/database/postgres/InsertSchemaPostgres.java
Updated schema conversion unit to handle REALs and BOOLEANs.

@@ -117,6 +117,8 @@ public class InsertSchemaPostgres {
                         .replace(" BIGINT", "")
                         .replace(" INT", "")
                         .replace(" TEXT", "")
+                        .replace(" REAL", "")
+                        .replace(" BOOLEAN", "")
                         .split(", ")) {
                     bw_labelProps.write(y);
                     bw_labelProps.newLine();
@@ -179,7 +181,9 @@ public class InsertSchemaPostgres {
                     .replace(" TEXT[]", "")
                     .replace(" BIGINT", "")
                     .replace(" TEXT", "")
-                    .replace(" INT", ""))
+                    .replace(" INT", "")
+                    .replace(" BOOLEAN", "")
+                    .replace(" REAL", ""))
                     .append(", ");
         }
 
@@ -384,6 +388,12 @@ public class InsertSchemaPostgres {
                 temp = value + ", ";
             } else if (inputField.endsWith("INT") && !inputField.contains("BIGINT")) {
                 int value = obj.get(inputField.split(" ")[0]).getAsInt();
+                temp = value + ", ";
+            } else if (inputField.endsWith("REAL") && !inputField.contains("BIGINT")) {
+                float value = obj.get(inputField.split(" ")[0]).getAsFloat();
+                temp = value + ", ";
+            } else if (inputField.endsWith("BOOLEAN") && !inputField.contains("BIGINT")) {
+                boolean value = obj.get(inputField.split(" ")[0]).getAsBoolean();
                 temp = value + ", ";
             } else if (inputField.endsWith("[]")) {
                 // is text with list property

### 2017-08-17 src/main/java/schema_conversion/PerformWork.java

@@ -225,6 +225,7 @@ class PerformWork implements Runnable {
         } else {
             // enter hacks here...
             if (key.equals("mono_time")) return "BIGINT";
+
             else {
                 String testValue = value.getAsString();
                 try {
@@ -232,7 +233,16 @@ class PerformWork implements Runnable {
                     Integer.parseInt(testValue);
                     return "INT";
                 } catch (NumberFormatException nfe) {
-                    return "TEXT";
+                    try {
+                        //noinspection ResultOfMethodCallIgnored
+                        Float.parseFloat(testValue);
+                        return "REAL";
+                    } catch (NumberFormatException nfe2) {
+                        boolean x = Boolean.parseBoolean(testValue);
+                        if (x) return "BOOLEAN";
+                        else if (testValue.equalsIgnoreCase("false")) return "BOOLEAN";
+                        else return "TEXT";
+                    }
                 }
             }
         }

### 2017-08-17 src/main/java/schema_conversion/SchemaConvert.java

@@ -190,7 +190,7 @@ public class SchemaConvert {
                 // escape character in SQL (' replaced with '')
                 line = line.replace("'", "''");
 
-                if (line.startsWith("create")) {
+                if (line.startsWith("create (")) {
                     if (firstLine) {
                         firstLine = false;
                     } else {

### 2017-08-17 src/main/java/translator/CypherTokenizer.java
Further bug fixes, updated README.

@@ -23,7 +23,6 @@ public class CypherTokenizer {
      * @param cypher      Cypher input.
      * @param DEBUG_PRINT Print out debug statements or not.
      * @return DecodedQuery object representing the Cypher input.
-     * @throws Exception Error somewhere.
      */
     public static DecodedQuery decode(String cypher, boolean DEBUG_PRINT) {
         return CypherTranslator.generateDecodedQuery(getTokenList(cypher, DEBUG_PRINT), cypherWalker);

### 2017-08-21 docs/latex/draftB.tex
Updated schema conversion unit to handle REALs and BOOLEANs.
Bug fixes and more testing and improved documentation.

@@
 
-\usepackage[hmargin={3 cm,3cm},top=1.5cm, marginpar=3cm]{geometry}
+\usepackage[hmargin={3cm,3cm},top=1.5cm, marginpar=3cm]{geometry}
 \usepackage{graphicx}
@@
 
-The codebase is entirely Java.
+The source code is entirely Java and is built using Maven.
 There are some additional features provided through the ANTLR tool, as described in \S \ref{sec:antlr}.
-There are some script/.bat files that help the tool execute SQL on Postgres and pipe the results back into the tool.
+There are some script/.bat files that help the tool execute SQL on Postgres and pipe the results back into the \emph{Cyp2SQL} tool.
 
 \subsection{Quick usage}
-Using the \texttt{cyp2sql-next.jar} file:
+Using the \texttt{cyp2sql-next.jar} file from the command line:
 \begin{verbatim}
@@
 \texttt{targetDatabaseName} & Name of the relational database for either the converted graph schema to be stored into, or for the translated queries to execute on. \\
-\texttt{dp \textbar{} -dn \textbar{} -r} & \textbf{Only needed when in query translation mode}.
+\texttt{dp \textbar{} -dn \textbar{} -r} & \textbf{Only needed when in query translation mode}. See \S \ref{ssec:usage} for further details.
 \end{tabular}
@@
 
-\begin{lstlisting}[language=Java, breaklines=true]
+\begin{lstlisting}[language=Java]
 import database.postgres.InsertSchemaPostgres;
@@
 boolean successConvert = SchemaConvert.translate(props);
-if (successConvert) InsertSchemaPostgres.executeSchemaChange(dbName, props);
+if (successConvert)
+InsertSchemaPostgres.executeSchemaChange(dbName, props);
 break;
@@
 
+\medskip
+
+The library and its dependencies can be added to a local Maven repository in the following way:
+
+\begin{lstlisting}[language=bash, breaklines=true, postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space}, backgroundcolor=\color{gray}, basicstyle=\color{white}\ttfamily]
+$ git clone git@gitlab.dtg.cl.cam.ac.uk:fresco-projects/cyp2sql-next.git
+$ cd cyp2sql-next
+$ mvn install
+\end{lstlisting}
+
 \section{Getting started}
@@
 
+The tool can currently handle the following datatypes: \texttt{TEXT}, \texttt{INT}, \texttt{REAL}, \texttt{BOOLEAN}, and \texttt{TEXT[]}.
+
 \subsection{Converting the graph database to a relational schema}
@@
 \subsection{Usage}
+\label{ssec:usage}
 The tool requires a running Neo4j instance to be up and running.

### 2017-08-21 src/main/java/intermediate_rep/CypOrder.java

@@ -4,7 +4,7 @@ package intermediate_rep;
  * Class for storing the order by clause of a cypher query (if there is one).
  */
 public class CypOrder {
-    private String nodeID;
+    private String idOrder;
     private String field;
     private String ascOrDesc;
 
@@ -21,7 +21,7 @@ public class CypOrder {
      *                  'asc', is recorded instead.
      */
     public CypOrder(String id, String field, String ascOrDesc) {
-        this.nodeID = id;
+        this.idOrder = id;
         this.field = field;
         this.ascOrDesc = ascOrDesc;
     }
@@ -36,7 +36,11 @@ public class CypOrder {
 
     @Override
     public String toString() {
-        return "(ID: " + this.nodeID + ", FIELD: " + this.field +
+        return "(ID: " + this.idOrder + ", FIELD: " + this.field +
                 ", ORDER_TYPE: " + this.ascOrDesc + ")";
+    }
+
+    public String getID() {
+        return idOrder;
     }
 }

### 2017-08-21 src/main/java/query_translation/sql/conversion_types/Multiple_With_Cypher.java

@@ -45,12 +45,19 @@ public class Multiple_With_Cypher extends AbstractConversion {
         String letter = String.valueOf(alphabet[numParts - 2]).toUpperCase();
         if (finalPartDQ.getMc().getRels().size() > 0) {
             int indexLastWhere = finalSQL.lastIndexOf("WHERE");
-            finalSQL = finalSQL.substring(0, indexLastWhere) + ", w" + letter + " WHERE " +
-                    finalSQL.substring(indexLastWhere + 6);
-            finalSQL = finalSQL.substring(0, finalSQL.length() - 1);
+
+            if (indexLastWhere == -1)
+                finalSQL = finalSQL.substring(0, finalSQL.length() - 1) + ", w" + letter + " WHERE ";
+            else {
+                finalSQL = finalSQL.substring(0, indexLastWhere) + ", w" + letter + " WHERE " +
+                        finalSQL.substring(indexLastWhere + 6);
+                finalSQL = finalSQL.substring(0, finalSQL.length() - 1);
+            }
+
             String correctPart = (mappingMultipleWith.keySet().contains(finalPartDQ.getMc().getNodes().get(0).getId()))
                     ? "a.a1" : "a.a2";
-            finalSQL = finalSQL + " AND " + correctPart + " = w" + letter + ".id";
+            finalSQL = (!finalSQL.endsWith(" WHERE ")) ? finalSQL + " AND " : finalSQL;
+            finalSQL = finalSQL + correctPart + " = w" + letter + ".id";
         } else finalSQL = finalSQL.replace(" nodes n01 ", " nodes n01, w" + letter + " ");
 
         multipleWithSQL.append(finalSQL).append(tempEnd);

### 2017-08-21 src/main/java/query_translation/sql/utilities_sql/MultipleRel.java

@@ -248,12 +248,17 @@ public class MultipleRel extends AbstractTranslation {
                                         .append(", ");
                             }
                         }
-                        if (cR.hasAggFunc() || cR.getCount() > 0) {
+                        if (cR.getCount() > 0) {
                             safSQL.setLength(safSQL.length() - 2);
                             safSQL.append(")");
                             safSQL.append(
                                     TranslateUtils.useAlias("count(" + cR.getNodeID() + ")", cR.getField(), alias))
                                     .append(", ");
+                        } else if (cR.hasAggFunc()) {
+                            safSQL.setLength(safSQL.length() - 2);
+                            safSQL.append(")");
+                            safSQL.append(
+                                    TranslateUtils.useAlias(cR.getNodeID(), cR.getField(), alias)).append(", ");
                         }
                         break;
                     }
@@ -286,21 +291,26 @@ public class MultipleRel extends AbstractTranslation {
                             safSQL.append(caseString).append(", ");
                         } else {
                             if (prop != null) {
-                                safSQL.append("n0").append(nodeTableCount).append(".").append(prop)
-                                        .append(TranslateUtils.useAlias(cR.getNodeID(), cR.getField(), alias))
-                                        .append(", ");
+                                safSQL.append("n0").append(nodeTableCount).append(".").append(prop);
+                                if (!cR.hasAggFunc())
+                                    safSQL.append(TranslateUtils.useAlias(cR.getNodeID(), cR.getField(), alias))
+                                            .append(", ");
                             } else {
                                 safSQL.append("n0").append(nodeTableCount)
                                         .append(".*").append(TranslateUtils.useAlias(cR.getNodeID(), cR.getField(), alias))
                                         .append(", ");
                             }
                         }
-                        if (cR.hasAggFunc() || cR.getCount() > 0) {
+                        if (cR.getCount() > 0) {
                             safSQL.setLength(safSQL.length() - 2);
                             safSQL.append(")");
                             safSQL.append(
                                     TranslateUtils.useAlias("count(" + cR.getNodeID() + ")", cR.getField(), alias))
                                     .append(", ");
+                        } else if (cR.hasAggFunc()) {
[…]

### 2017-08-21 src/main/java/query_translation/sql/utilities_sql/SQLTranslate.java

@@ -45,7 +45,8 @@ public class SQLTranslate {
         } else {
             MultipleRel mr = new MultipleRel();
             sql = mr.translate(sql, decodedQuery, props);
-            if (decodedQuery.getCypherAdditionalInfo().hasCount() && decodedQuery.getRc().getItems().size() > 1)
+            if ((decodedQuery.getCypherAdditionalInfo().hasCount() || decodedQuery.getCypherAdditionalInfo().hasAgg())
+                    && decodedQuery.getRc().getItems().size() > 1)
                 sql = obtainGroupByClause(decodedQuery.getRc(), sql, props);
         }
 
@@ -216,13 +217,22 @@ public class SQLTranslate {
         }
 
         for (CypOrder cO : orderC.getItems()) {
+            int posInReturn = 1;
+            for (CypReturn cR : rc.getItems()) {
+                if (cR.getType().equals("node") && cR.getNodeID().equals(cO.getID())) {
+                    nodeID = "n0" + posInReturn;
+                    break;
+                }
+                if (cR.getType().equals("node")) posInReturn++;
+            }
+
             if (cO.getField().startsWith("count")) {
                 sql.append("count(").append(nodeID).append(")").append(cO.getAscOrDesc()).append(", ");
                 break;
             }
-            sql.append(nodeID).append(".").append(cO.getField()).append(" ")
-                    .append(cO.getAscOrDesc());
-            sql.append(", ");
+
+            sql.append(nodeID).append(".")
+                    .append(cO.getField()).append(" ").append(cO.getAscOrDesc()).append(", ");
         }
 
         sql.setLength(sql.length() - 2);

### 2017-08-21 src/main/java/query_translation/sql/utilities_sql/TranslateUtils.java

@@ -209,7 +209,10 @@ class TranslateUtils {
             else if (list) sql.append(v).append(" ");
             else if (v.startsWith("id(")) {
                 String id = v.substring(v.indexOf("(") + 1, v.length() - 1);
-                String withTable = Multiple_With_Cypher.mappingMultipleWith.get(id);
+                String withTable;
+                if (Multiple_With_Cypher.mappingMultipleWith != null)
+                    withTable = Multiple_With_Cypher.mappingMultipleWith.get(id);
+                else withTable = "n2";
                 sql.append(withTable).append(".id");
             } else if (v.contains(".")) {
                 String idAndValue[] = v.split("\\.");
@@ -316,6 +319,7 @@ class TranslateUtils {
                 // TODO: extend for more of the aggregate functions.
                 if (key.startsWith("collect")) key = key.substring(8, key.length() - 1);
                 if (key.startsWith("sum")) key = key.substring(4, key.length() - 1);
+                if (key.startsWith("max")) key = key.substring(4, key.length() - 1);
                 if (field != null) {
                     if (key.equals((nodeID) + "." + (field))) {
                         return (" AS " + alias.get(s));

### 2017-08-21 src/main/java/translator/CypherWalker.java

@@ -253,4 +253,8 @@ public class CypherWalker extends CypherBaseListener {
     boolean hasMax() {
         return hasMax;
     }
+
+    public boolean hasAgg() {
+        return (hasAverage || hasSum || hasMin || hasMax);
+    }
 }

### 2017-08-21 src/test/java/production/C2SBenchmarkTest.java

@@ -28,7 +28,7 @@ public class C2SBenchmarkTest {
      */
     @Test
     public void allQueriesTest() throws Exception {
-        boolean run = false;
+        boolean run = true;
         if (run) {
             // Using JUnit, two files are created in temp storage to hold the results outputted from both databases.
             File f_neo4j = folder.newFile("neo4j_results.txt");

### 2017-08-22 docs/latex/draftB.tex
Bug fixes and more testing and improved documentation.

@@
 \label{ssec:grreltheory}
-All the nodes from the graph are stored in one relation (nodes), and all the relationships are stored in a separate relation (edges).
+All the nodes from the graph are stored in one relation (\emph{Nodes}), and all the relationships are stored in a separate relation (\emph{Edges}).
 
@@
 For graph functions to be computable on the relational side, the schema converter also builds two materialised views –- \texttt{adjlist\_from} \& \texttt{adjlist\_to}.
+For the performance of the queries to be better when relationships are used, a unique index constraint is added to the \emph{Edges} relation as well.
+It is recommended to add even further indices where applicable for larger boosts in speed.
 

### 2017-08-22 src/main/java/database/postgres/InsertSchemaPostgres.java

@@ -43,6 +43,7 @@ public class InsertSchemaPostgres {
             PostgresDriver.createInsert(PostgresConstants.CYPHER_ITERATE);
             PostgresDriver.createInsert(PostgresConstants.UNIQUE_ARR_FUNC);
             PostgresDriver.createInsert(PostgresConstants.AUTO_SEQ_QUERY);
+            PostgresDriver.createInsert(PostgresConstants.EDGES_INDEX);
         } catch (SQLException e) {
             e.printStackTrace();
         }

### 2017-08-22 src/main/java/database/postgres/PostgresConstants.java

@@ -57,4 +57,7 @@ class PostgresConstants {
     static final String AUTO_SEQ_QUERY = "CREATE SEQUENCE nodes_id_seq;\n" +
             "ALTER TABLE nodes ALTER id SET DEFAULT NEXTVAL('nodes_id_seq');";
 
+    // indexes to improve the performance (particularly of larger databases).
+    static final String EDGES_INDEX = "CREATE UNIQUE INDEX edges_uniq_1 ON edges(idl, idr, type);";
+
 }

### 2017-08-22 src/main/java/database/Neo4jDriver.java
Some bug fixes to ORDER BY and GROUP BY parts. Schema conversion tested against additional graph databases.

@@ -78,10 +78,14 @@ public class Neo4jDriver {
                                 }
                             } else {
                                 if (props.getListFields().contains(key)) {
-                                    writer.println(key + " : {" + entry.getValue()
-                                            .toString()
-                                            .replace("[", "")
-                                            .replace("]", "") + "}");
+                                    try {
+                                        writer.println(key + " : {" + entry.getValue()
+                                                .toString()
+                                                .replace("[", "")
+                                                .replace("]", "") + "}");
+                                    } catch (NullPointerException npe) {
+                                        // do nothing
+                                    }
                                 } else {
                                     if (key.startsWith("id(")) key = "id";
                                     writer.println(key + " : " + entry.getValue());

### 2017-08-22 src/main/java/database/postgres/InsertSchemaPostgres.java

@@ -43,13 +43,16 @@ public class InsertSchemaPostgres {
             PostgresDriver.createInsert(PostgresConstants.CYPHER_ITERATE);
             PostgresDriver.createInsert(PostgresConstants.UNIQUE_ARR_FUNC);
             PostgresDriver.createInsert(PostgresConstants.AUTO_SEQ_QUERY);
+
+            addFieldsToMetaFile(props);
+
+            // potentially risky insert, presumes no duplicated edges
             PostgresDriver.createInsert(PostgresConstants.EDGES_INDEX);
         } catch (SQLException e) {
             e.printStackTrace();
+        } finally {
+            PostgresDriver.closeConnection();
         }
-
-        addFieldsToMetaFile(props);
-        PostgresDriver.closeConnection();
     }
 
     /**

### 2017-08-22 src/main/java/database/postgres/PostgresDriver.java

@@ -86,7 +86,7 @@ public class PostgresDriver {
      * @param props       C2SProperties object (should already be initialised).
      * @throws SQLException Thrown if there is an error in the SQL statement.
      */
-    public static void select(String query, String database, File pg_results, boolean printOutput, C2SProperties props)
+    public static int select(String query, String database, File pg_results, boolean printOutput, C2SProperties props)
             throws SQLException {
         if (!DB_OPEN) PostgresDriver.createConnection(database, props);
         Statement stmt;
@@ -124,6 +124,7 @@ public class PostgresDriver {
         C2SMain.needToPrintID = false;
         C2SMain.numResultsPostgres = numRecords;
         PostgresDriver.closeConnection();
+        return numRecords;
     }
 
     /**

### 2017-08-22 src/main/java/production/C2SProperties.java

@@ -12,6 +12,10 @@ import java.util.Properties;
  * documentation for the getLocalProperties() method for more detailed information.
  */
 public class C2SProperties {
+    // Some Neo4j fields need to be adapted to handle lists. Currently, this requires the
+    // user to manually outline the fields that *may* contain lists. These are stored
+    // in this array (they are read from a file titled lists.txt).
+    public static ArrayList<String> listFields;
     private String propsLocation;
     private String neo4jSchema;
     private String wspace;
@@ -21,11 +25,6 @@ public class C2SProperties {
     private String neoPW;
     private String postUN;
     private String postPW;
-
-    // Some Neo4j fields need to be adapted to handle lists. Currently, this requires the
-    // user to manually outline the fields that *may* contain lists. These are stored
-    // in this array (they are read from a file titled lists.txt).
-    private ArrayList<String> listFields;
 
     /**
      * Constructor method. Sets the location of the properties file (c2s_props.properties)
@@ -88,7 +87,7 @@ public class C2SProperties {
         this.neo4jRes = neo4jRes;
     }
 
-    String getSqlRes() {
+    public String getSqlRes() {
         return sqlRes;
     }
 
@@ -161,7 +160,7 @@ public class C2SProperties {
         return listFields;
     }
 
-    private void setListFields(ArrayList<String> listFields) {
-        this.listFields = listFields;
+    private void setListFields(ArrayList<String> listFieldsZ) {
+        listFields = listFieldsZ;
     }
 }

### 2017-08-22 src/main/java/query_translation/sql/utilities_sql/MultipleRel.java

@@ -337,26 +337,26 @@ public class MultipleRel extends AbstractTranslation {
                                 safSQL.append(caseString).append(", ");
                             } else {
                                 if (field != null) {
-                                    safSQL.append(idRel).append(".").append(field)
-                                            .append(TranslateUtils.useAlias(cR.getNodeID(), cR.getField(), alias))
-                                            .append(", ");
+                                    safSQL.append(idRel).append(".").append(field);
+                                    if (!cR.hasAggFunc())
+                                        safSQL.append(TranslateUtils.useAlias(cR.getNodeID(), cR.getField(), alias))
+                                                .append(", ");
                                 } else {
                                     safSQL.append(idRel).append(".*")
                                             .append(TranslateUtils.useAlias(cR.getNodeID(), cR.getField(), alias))
                                             .append(", ");
                                 }
                             }
-                            if (cR.hasAggFunc() || cR.getCount() > 0) {
+                            if (cR.getCount() > 0) {
                                 safSQL.setLength(safSQL.length() - 2);
                                 safSQL.append(")");
-                                if (cR.hasAggFunc())
-                                    safSQL.append(
-                                            TranslateUtils.useAlias(
-                                                    "collect(" + cR.getNodeID() + ")", null, alias))
-                                            .append(", ");
-                                else safSQL.append(
-                                        TranslateUtils.useAlias("count(" + cR.getNodeID() + ")", null, alias))
+                                safSQL.append(
+                                        TranslateUtils.useAlias("count(" + cR.getNodeID() + ")", cR.getField(), alias))
                                         .append(", ");
+                            } else if (cR.hasAggFunc()) {
+                                safSQL.append(")");
+                                safSQL.append(
+                                        TranslateUtils.useAlias(cR.getNodeID(), cR.getField(), alias)).append(", ");
                             }
                             break;
                         }
@@ -455,6 +455,8 @@ public class MultipleRel extends AbstractTranslation {
                 switch (cR.getType()) {
                     case "node":
                         if (!(cR.getCount() > 0 && returnC.getItems().size() > 1)) {
+                            if (WithSQL.withMapping.containsKey(cR.getNodeID())) break;
+
                             if (!nodesSeenSoFar.contains(cR.getNodeID())) {
[…]

### 2017-08-22 src/main/java/query_translation/sql/utilities_sql/SQLTranslate.java

@@ -217,22 +217,25 @@ public class SQLTranslate {
         }
 
         for (CypOrder cO : orderC.getItems()) {
-            int posInReturn = 1;
-            for (CypReturn cR : rc.getItems()) {
-                if (cR.getType().equals("node") && cR.getNodeID().equals(cO.getID())) {
-                    nodeID = "n0" + posInReturn;
+            if (cO.getID() == null) sql.append(cO.getField()).append(" ").append(cO.getAscOrDesc()).append(", ");
+            else {
+                int posInReturn = 1;
+                for (CypReturn cR : rc.getItems()) {
+                    if (cR.getType().equals("node") && cR.getNodeID().equals(cO.getID())) {
+                        nodeID = "n0" + posInReturn;
+                        break;
+                    }
+                    if (cR.getType().equals("node")) posInReturn++;
+                }
+
+                if (cO.getField().startsWith("count")) {
+                    sql.append("count(").append(nodeID).append(")").append(cO.getAscOrDesc()).append(", ");
                     break;
                 }
-                if (cR.getType().equals("node")) posInReturn++;
-            }
 
-            if (cO.getField().startsWith("count")) {
-                sql.append("count(").append(nodeID).append(")").append(cO.getAscOrDesc()).append(", ");
-                break;
+                sql.append(nodeID).append(".")
+                        .append(cO.getField()).append(" ").append(cO.getAscOrDesc()).append(", ");
             }
-
-            sql.append(nodeID).append(".")
-                    .append(cO.getField()).append(" ").append(cO.getAscOrDesc()).append(", ");
         }
 
         sql.setLength(sql.length() - 2);
@@ -262,9 +265,12 @@ public class SQLTranslate {
                 nodeTableCount++;
             }
 
-            if (cR.getField() != null && cR.getCount() == COUNT_FALSE) {
-                sql.append("n0").append(nodeTableCount).append(".").append(cR.getField()).append(", ");
-            } else if (cR.getCount() == COUNT_FALSE) {
+            if (cR.getField() != null && cR.getCount() == COUNT_FALSE && (cR.hasAggFunc() && cR.getAggFunc() == 1)) {
+                String ident;
+                if (cR.getType().equals("node")) ident = "n0" + nodeTableCount;
+                else ident = "a";
+                sql.append(ident).append(".").append(cR.getField()).append(", ");
+            } else if (cR.getCount() == COUNT_FALSE && !cR.hasAggFunc()) {
[…]

### 2017-08-22 src/main/java/query_translation/sql/utilities_sql/SingleVar.java

@@ -55,11 +55,12 @@ public class SingleVar extends AbstractTranslation {
 
         if (cn2.getType() != null && table.equals("nodes")) {
             if (!hasWhere) {
-                sql.append(" WHERE n01.label LIKE");
+                sql.append(" WHERE n01.label LIKE ");
             } else {
-                sql.append(" AND n01.label LIKE");
+                if (!sql.toString().toLowerCase().endsWith(" and ")) sql.append(" AND");
+                sql.append(" n01.label LIKE ");
             }
-            sql.append(" ").append(TranslateUtils.genLabelLike(cn2, "n01"));
+            sql.append(TranslateUtils.genLabelLike(cn2, "n01"));
         }
 
         return sql;

### 2017-08-22 src/main/java/query_translation/sql/utilities_sql/TranslateUtils.java

@@ -34,7 +34,7 @@ class TranslateUtils {
             String value;
             if (entry.getValue().isJsonArray()) {
                 value = "ARRAY" + entry.getValue().getAsJsonArray().toString();
-            } else if (entry.getKey().equals("name")) {
+            } else if (C2SProperties.listFields.contains(entry.getKey())) {
                 value = "ARRAY[" + entry.getValue().toString() + "]";
             } else value = entry.getValue().getAsString();
 

### 2017-08-22 src/main/java/schema_conversion/PerformWork.java

@@ -132,6 +132,8 @@ class PerformWork implements Runnable {
             if (!SchemaConvert.edgesRelLabels.contains(entry.getKey() + " TEXT") &&
                     !SchemaConvert.edgesRelLabels.contains(entry.getKey() + " INT") &&
                     !SchemaConvert.edgesRelLabels.contains(entry.getKey() + " BIGINT") &&
+                    !SchemaConvert.edgesRelLabels.contains(entry.getKey() + " REAL") &&
+                    !SchemaConvert.edgesRelLabels.contains(entry.getKey() + " BOOLEAN") &&
                     !SchemaConvert.edgesRelLabels.contains(entry.getKey() + " TEXT[]")) {
                 String textToAdd = entry.getKey() + " " + type;
                 if (propsX.getListFields().contains(entry.getKey()))
@@ -195,6 +197,8 @@ class PerformWork implements Runnable {
 
             if (!SchemaConvert.nodeRelLabels.contains(entry.getKey() + " TEXT") &&
                     !SchemaConvert.nodeRelLabels.contains(entry.getKey() + " INT") &&
+                    !SchemaConvert.nodeRelLabels.contains(entry.getKey() + " REAL") &&
+                    !SchemaConvert.nodeRelLabels.contains(entry.getKey() + " BOOLEAN") &&
                     !SchemaConvert.nodeRelLabels.contains(entry.getKey() + " BIGINT") &&
                     !SchemaConvert.nodeRelLabels.contains(entry.getKey() + " TEXT[]")) {
                 SchemaConvert.nodeRelLabels.add(entry.getKey() + " " + type);
@@ -211,7 +215,7 @@ class PerformWork implements Runnable {
 
     /**
      * Calculate the datatype of a key, based on an example value presented to this method. This method mostly
-     * works as expected, however fails in some cases where it inteprets a field as an INT, where in fact it
+     * works as expected, however fails in some cases where it interprets a field as an INT, where in fact it
      * should be of type BIGINT. The tool cannot currently handle this automatically, however can be fixed
      * manually in the code.
      *

### 2017-08-22 src/main/java/translator/CypherTranslator.java

@@ -95,7 +95,7 @@ class CypherTranslator {
                 // if ORDER BY is present in the query
                 OrderClause orderC = null;
                 if (orderClause != null) {
-                    orderC = orderDecode(orderClause);
+                    orderC = orderDecode(orderClause, cypWalker);
                 }
 
                 int skipAmount = (posOfSkip != -1) ? cypWalker.getSkipAmount() : -1;
@@ -633,9 +633,10 @@ class CypherTranslator {
      * Translate the ORDER BY part of the Cypher input (if one is present).
      *
      * @param orderClause Tokenised form of the ORDER BY part of the Cypher input.
+     * @param cypWalker
      * @return OrderClause representing the tokenised input.
      */
-    private static OrderClause orderDecode(List<String> orderClause) throws DQInvalidException {
+    private static OrderClause orderDecode(List<String> orderClause, CypherWalker cypWalker) throws DQInvalidException {
         OrderClause o = new OrderClause();
 
         List<CypOrder> items = new ArrayList<>();
@@ -647,11 +648,11 @@ class CypherTranslator {
             int posComma = orderClause.indexOf(",");
             currentWorking = orderClause.subList(0, posComma);
             orderClause = orderClause.subList(posComma + 1, orderClause.size());
-            items.add(extractOrder(currentWorking));
+            items.add(extractOrder(currentWorking, cypWalker));
         }
 
         if (!orderClause.isEmpty()) {
-            items.add(extractOrder(orderClause));
+            items.add(extractOrder(orderClause, cypWalker));
         }
 
         o.setItems(items);
@@ -661,10 +662,11 @@ class CypherTranslator {
     /**
      * Extracts a CypOrder object from the tokenised input.
      *
-     * @param clause Token list describing the ORDER BY part of the original Cypher query.
+     * @param clause    Token list describing the ORDER BY part of the original Cypher query.
+     * @param cypWalker
      * @return CypOrder object matching the inputted token list.
      */
-    private static CypOrder extractOrder(List<String> clause) throws DQInvalidException {
+    private static CypOrder extractOrder(List<String> clause, CypherWalker cypWalker) throws DQInvalidException {
         /*
         In order below of the types of ORDER BY structures that can be successfully parsed, and the
         string representation of what they are parsed into when they become CypOrder objects.
@@ -675,6 +677,7 @@ class CypherTranslator {
[…]

### 2017-08-22 src/main/java/translator/CypherWalker.java

@@ -210,6 +210,10 @@ public class CypherWalker extends CypherBaseListener {
         return whereClause;
     }
 
+    String getOrderClause() {
+        return orderClause;
+    }
+
     public boolean hasDistinct() {
         return hasDistinct;
     }

### 2017-08-24 docs/latex/draftB.tex
Some bug fixes to ORDER BY and GROUP BY parts. Schema conversion tested against additional graph databases.
Licensing added.

@@
 
-\subsection{Known issues}
+\section{Known issues}
 \begin{itemize}
 \item The use of \texttt{WITH} is still quite constrained in its use
+\item Alias names cannot be the same as field names -- for example, \texttt{a.points AS points} would not work; \texttt{a.points AS points\_x} is fine however
+\item Some labels from Neo4j may map to reserved keywords in SQL (such as \emph{Group} and \emph{User}) - in these cases the optimised table for these labels cannot be created
 \item When predicates in the \texttt{WHERE} clause are added, they may sometimes lead to incorrect translations (mostly in the \texttt{WITH} parts of a multiple relationship Cypher input).
@@
 
-\subsection{List of test queries from OPUS}
+\section{List of test queries from OPUS}
 \begin{verbatim}

### 2017-08-24 src/main/java/database/Neo4jDriver.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package database;
 
 import org.neo4j.driver.v1.*;

### 2017-08-24 src/main/java/database/SchemaConstants.java

@@ -17,10 +17,17 @@
  * limitations under the License.
  */
 
-package exceptions;
+package database;
 
-public class DQInvalidException extends Exception {
-    public DQInvalidException(String s) {
-        super(s);
-    }
+import java.util.Arrays;
+import java.util.List;
+
+public class SchemaConstants {
+    // notice leading space!
+    public static List<String> DATATYPES =
+            Arrays.asList(" TEXT[]", " BIGINT", " TEXT", " INT", " BOOLEAN", " REAL");
+
+    // reserved keywords in SQL that therefore cannot be used as relation names!
+    public static List<String> RESERVED_KW =
+            Arrays.asList("group", "user");
 }

### 2017-08-24 src/main/java/database/key_value_hazelcast/KeyValueTest.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package database.key_value_hazelcast;
 
 import com.google.gson.JsonObject;

### 2017-08-24 src/main/java/database/postgres/InsertSchemaPostgres.java

@@ -1,8 +1,28 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package database.postgres;
 
 import com.google.gson.JsonArray;
 import com.google.gson.JsonObject;
 import com.google.gson.JsonParser;
+import database.SchemaConstants;
 import production.C2SProperties;
 import schema_conversion.SchemaConvert;
 
@@ -49,7 +69,10 @@ public class InsertSchemaPostgres {
             // potentially risky insert, presumes no duplicated edges
             PostgresDriver.createInsert(PostgresConstants.EDGES_INDEX);
         } catch (SQLException e) {
-            e.printStackTrace();
+            if (e.getMessage().contains("not create unique index")) {
+                System.err.println("Could not create a unique index in the edges relation " +
+                        "as there is a duplicated relationship present.");
+            } else e.printStackTrace();
         } finally {
             PostgresDriver.closeConnection();
         }
@@ -110,7 +133,7 @@ public class InsertSchemaPostgres {
 
             for (String label : SchemaConvert.labelMappings.keySet()) {
                 String tableLabel = label.replace(", ", "_");
-                if (!tableLabel.equals("group")) {
+                if (!SchemaConstants.RESERVED_KW.contains(tableLabel)) {
                     sb.append("CREATE TABLE ").append(tableLabel).append("(");
                     sb.append(SchemaConvert.labelMappings.get(label));
                     sb.append("); ");
@@ -180,18 +203,15 @@ public class InsertSchemaPostgres {
      */
     private static StringBuilder insertDataForLabels(StringBuilder sb, String label, JsonObject o) {
         String tableLabel = label.replace(", ", "_");
-        if (!tableLabel.equals("group")) {
+        if (!SchemaConstants.RESERVED_KW.contains(tableLabel)) {
[…]

### 2017-08-24 src/main/java/database/postgres/PostgresConstants.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package database.postgres;
 
 /**

### 2017-08-24 src/main/java/database/postgres/PostgresDriver.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package database.postgres;
 
 import production.C2SMain;

### 2017-08-24 src/main/java/exceptions/ConversionSQLException.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package exceptions;
 
 public class ConversionSQLException extends Exception {

### 2017-08-24 src/main/java/intermediate_rep/CypForEach.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package intermediate_rep;
 
 import java.util.HashMap;

### 2017-08-24 src/main/java/intermediate_rep/CypOrder.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package intermediate_rep;
 
 /**

### 2017-08-24 src/main/java/intermediate_rep/CypRel.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package intermediate_rep;
 
 import com.google.gson.JsonObject;

### 2017-08-24 src/main/java/intermediate_rep/CypReturn.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package intermediate_rep;
 
 import exceptions.DQInvalidException;

### 2017-08-24 src/main/java/intermediate_rep/DecodedQuery.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package intermediate_rep;
 
 import translator.CypherWalker;

### 2017-08-24 src/main/java/intermediate_rep/MatchClause.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package intermediate_rep;
 
 import java.util.ArrayList;

### 2017-08-24 src/main/java/intermediate_rep/ReturnClause.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package intermediate_rep;
 
 import java.util.List;

### 2017-08-24 src/main/java/parsing_lexing/CypherBaseListener.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package parsing_lexing;// Generated from Cypher.g4 by ANTLR 4.7
 
 import org.antlr.v4.runtime.ParserRuleContext;
@@ -16,1218 +35,1421 @@ public class CypherBaseListener implements CypherListener {
      * <p>The default implementation does nothing.</p>
      */
     @Override public void enterCypher(CypherParser.CypherContext ctx) { }
+
     /**
      * {@inheritDoc}
      *
      * <p>The default implementation does nothing.</p>
      */
     @Override public void exitCypher(CypherParser.CypherContext ctx) { }
+
     /**
      * {@inheritDoc}
      *
      * <p>The default implementation does nothing.</p>
      */
     @Override public void enterStatement(CypherParser.StatementContext ctx) { }
+
     /**
      * {@inheritDoc}
      *
      * <p>The default implementation does nothing.</p>
      */
     @Override public void exitStatement(CypherParser.StatementContext ctx) { }
+
     /**
      * {@inheritDoc}
      *
      * <p>The default implementation does nothing.</p>
      */
     @Override public void enterQuery(CypherParser.QueryContext ctx) { }
+
     /**
      * {@inheritDoc}
      *
      * <p>The default implementation does nothing.</p>
      */
     @Override public void exitQuery(CypherParser.QueryContext ctx) { }
+
     /**
      * {@inheritDoc}
      *
      * <p>The default implementation does nothing.</p>
      */
     @Override public void enterRegularQuery(CypherParser.RegularQueryContext ctx) { }
+
     /**
      * {@inheritDoc}
      *
      * <p>The default implementation does nothing.</p>
      */
[…]

### 2017-08-24 src/main/java/parsing_lexing/CypherLexer.java

@@ -1,12 +1,30 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package parsing_lexing;// Generated from Cypher.g4 by ANTLR 4.7
-import org.antlr.v4.runtime.Lexer;
-import org.antlr.v4.runtime.CharStream;
-import org.antlr.v4.runtime.Token;
-import org.antlr.v4.runtime.TokenStream;
+
 import org.antlr.v4.runtime.*;
-import org.antlr.v4.runtime.atn.*;
+import org.antlr.v4.runtime.atn.ATN;
+import org.antlr.v4.runtime.atn.ATNDeserializer;
+import org.antlr.v4.runtime.atn.LexerATNSimulator;
+import org.antlr.v4.runtime.atn.PredictionContextCache;
 import org.antlr.v4.runtime.dfa.DFA;
-import org.antlr.v4.runtime.misc.*;
 
 @SuppressWarnings({"all", "warnings", "unchecked", "unused", "cast"})
 public class CypherLexer extends Lexer {
@@ -107,6 +125,7 @@ public class CypherLexer extends Lexer {
      */
     @Deprecated
     public static final String[] tokenNames;
+
     static {
         tokenNames = new String[_SYMBOLIC_NAMES.length];
         for (int i = 0; i < tokenNames.length; i++) {
@@ -703,6 +722,7 @@ public class CypherLexer extends Lexer {
                     "\u0418\u041a\2";
     public static final ATN _ATN =
             new ATNDeserializer().deserialize(_serializedATN.toCharArray());
+
     static {
         _decisionToDFA = new DFA[_ATN.getNumberOfDecisions()];
         for (int i = 0; i < _ATN.getNumberOfDecisions(); i++) {

### 2017-08-24 src/main/java/parsing_lexing/CypherListener.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package parsing_lexing;// Generated from Cypher.g4 by ANTLR 4.7
 import org.antlr.v4.runtime.tree.ParseTreeListener;
 
@@ -6,1024 +25,1227 @@ import org.antlr.v4.runtime.tree.ParseTreeListener;
  * {@link CypherParser}.
  */
 public interface CypherListener extends ParseTreeListener {
-	/**
-	 * Enter a parse tree produced by {@link CypherParser#cypher}.
-	 * @param ctx the parse tree
-	 */
-	void enterCypher(CypherParser.CypherContext ctx);
-	/**
-	 * Exit a parse tree produced by {@link CypherParser#cypher}.
-	 * @param ctx the parse tree
-	 */
-	void exitCypher(CypherParser.CypherContext ctx);
-	/**
-	 * Enter a parse tree produced by {@link CypherParser#statement}.
-	 * @param ctx the parse tree
-	 */
-	void enterStatement(CypherParser.StatementContext ctx);
-	/**
-	 * Exit a parse tree produced by {@link CypherParser#statement}.
-	 * @param ctx the parse tree
-	 */
-	void exitStatement(CypherParser.StatementContext ctx);
-	/**
-	 * Enter a parse tree produced by {@link CypherParser#query}.
-	 * @param ctx the parse tree
-	 */
-	void enterQuery(CypherParser.QueryContext ctx);
-	/**
-	 * Exit a parse tree produced by {@link CypherParser#query}.
-	 * @param ctx the parse tree
-	 */
[…]

### 2017-08-24 src/main/java/production/C2SBenchmark.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package production;
 
 import org.apache.commons.io.FileUtils;

### 2017-08-24 src/main/java/production/C2SInteractive.java

@@ -1,6 +1,26 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package production;
 
 import jline.TerminalFactory;
+import jline.UnsupportedTerminal;
 import jline.console.ConsoleReader;
 
 import java.io.File;
@@ -24,6 +44,8 @@ class C2SInteractive {
         System.out.println("PRINT TO FILE : " + ((C2SMain.printBool) ? "enabled" : "disabled"));
         System.out.println("Cypher to SQL Translator Tool v1.1");
         System.out.println("To exit, type :exit.");
+
+        jline.TerminalFactory.registerFlavor(jline.TerminalFactory.Flavor.WINDOWS, UnsupportedTerminal.class);
 
         try {
             ConsoleReader console = new ConsoleReader();

### 2017-08-24 src/main/java/production/C2SMain.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package production;
 
 import database.Neo4jDriver;

### 2017-08-24 src/main/java/production/C2SProperties.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package production;
 
 import java.io.BufferedReader;

### 2017-08-24 src/main/java/query_translation/sql/conversion_types/ForEach_Cypher.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package query_translation.sql.conversion_types;
 
 import exceptions.DQInvalidException;

### 2017-08-24 src/main/java/query_translation/sql/utilities_sql/AbstractTranslation.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package query_translation.sql.utilities_sql;
 
 import exceptions.DQInvalidException;

### 2017-08-24 src/main/java/query_translation/sql/utilities_sql/ForEach.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package query_translation.sql.utilities_sql;
 
 import intermediate_rep.CypForEach;

### 2017-08-24 src/main/java/query_translation/sql/utilities_sql/InsertUtils.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package query_translation.sql.utilities_sql;
 
 import com.google.gson.JsonElement;

### 2017-08-24 src/main/java/query_translation/sql/utilities_sql/IterateSQL.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package query_translation.sql.utilities_sql;
 
 import com.google.gson.JsonObject;

### 2017-08-24 src/main/java/query_translation/sql/utilities_sql/MultipleRel.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package query_translation.sql.utilities_sql;
 
 import com.google.gson.JsonObject;
@@ -542,8 +561,8 @@ public class MultipleRel extends AbstractTranslation {
                         int pos = cN.getPosInClause();
                         if (pos == 1) where.append(" AND wA.id = a.a1");
                         else {
-                            where.append(" AND wA.id = ").append(alphabet[pos - 1])
-                                    .append(".").append(alphabet[pos - 1]).append(1);
+                            where.append(" AND wA.id = ").append(alphabet[pos - 2])
+                                    .append(".").append(alphabet[pos - 2]).append(2);
                         }
                     }
                 }

### 2017-08-24 src/main/java/query_translation/sql/utilities_sql/SQLTranslate.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package query_translation.sql.utilities_sql;
 
 import exceptions.DQInvalidException;
@@ -24,6 +43,8 @@ import static intermediate_rep.CypCount.COUNT_FALSE;
  * Agnostic to the methods above is appending the ORDER BY, GROUP BY, LIMIT and SKIP elements.
  */
 public class SQLTranslate {
+    private static final char[] alphabet = "abcdefghijklmnopqrstuvwxyz".toCharArray();
+
     public static String translateRead(DecodedQuery decodedQuery, C2SProperties props)
             throws DQInvalidException, IOException {
         // SQL built up from a StringBuilder object.
@@ -223,6 +244,10 @@ public class SQLTranslate {
                 for (CypReturn cR : rc.getItems()) {
                     if (cR.getType().equals("node") && cR.getNodeID().equals(cO.getID())) {
                         nodeID = "n0" + posInReturn;
+                        break;
+                    }
+                    if (cR.getType().equals("rel") && cR.getNodeID().equals(cO.getID())) {
+                        nodeID = String.valueOf(alphabet[cR.getPosInClause() - 1]);
                         break;
                     }
                     if (cR.getType().equals("node")) posInReturn++;

### 2017-08-24 src/main/java/query_translation/sql/utilities_sql/SingleVar.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package query_translation.sql.utilities_sql;
 
 import intermediate_rep.*;

### 2017-08-24 src/main/java/query_translation/sql/utilities_sql/UnionSQL.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package query_translation.sql.utilities_sql;
 
 import java.util.ArrayList;

### 2017-08-24 src/main/java/schema_conversion/PerformWork.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package schema_conversion;
 
 import com.google.gson.JsonArray;

### 2017-08-24 src/main/java/schema_conversion/SchemaConvert.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package schema_conversion;
 
 import com.google.gson.JsonParser;

### 2017-08-24 src/main/java/translator/CypherTokenizer.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package translator;
 
 import intermediate_rep.DecodedQuery;

### 2017-08-24 src/main/java/translator/CypherTranslator.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package translator;
 
 import com.google.gson.JsonObject;
@@ -255,7 +274,7 @@ class CypherTranslator {
      * @param clause Token version of the clause.
      * @param m      Internal representation of the MATCH clause.
      * @return An ArrayList of CypRel objects that correspond to the relationships (if any) described by Cypher.
-     * @throws Exception Error in the clause.
+     * @throws DQInvalidException Relationship structure invalid in some way.
      */
     private static ArrayList<CypRel> extractRels(List<String> clause, MatchClause m) throws DQInvalidException {
         ArrayList<CypRel> rels = new ArrayList<>();
@@ -523,7 +542,7 @@ class CypherTranslator {
      *                     the RETURN part can be).
      * @param cypWalker    Additional information about the query.
      * @return ReturnClause object
-     * @throws Exception Error in creating the ReturnClause object (exception thrown in extractReturn method).
+     * @throws DQInvalidException Error in creating the ReturnClause object (exception thrown in extractReturn method).
      */
     private static ReturnClause returnDecode(List<String> returnClause, MatchClause matchC, CypherWalker cypWalker)
             throws DQInvalidException {
@@ -663,7 +682,7 @@ class CypherTranslator {
      * Extracts a CypOrder object from the tokenised input.
      *
      * @param clause    Token list describing the ORDER BY part of the original Cypher query.
-     * @param cypWalker
+     * @param cypWalker CypherWalker object.
      * @return CypOrder object matching the inputted token list.
      */
     private static CypOrder extractOrder(List<String> clause, CypherWalker cypWalker) throws DQInvalidException {

### 2017-08-24 src/main/java/translator/CypherWalker.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package translator;
 
 import parsing_lexing.CypherBaseListener;

### 2017-08-24 src/test/java/production/C2SBenchmarkTest.java

@@ -1,3 +1,22 @@
+/*
+ * Copyright (c) 2017.
+ *
+ * Oliver Crawford <o.crawford@hotmail.co.uk>
+ * Lucian Carata <lc525@cam.ac.uk>
+ *
+ * Licensed under the Apache License, Version 2.0 (the "License");
+ * you may not use this file except in compliance with the License.
+ * You may obtain a copy of the License at
+ *
+ *     http://www.apache.org/licenses/LICENSE-2.0
+ *
+ * Unless required by applicable law or agreed to in writing, software
+ * distributed under the License is distributed on an "AS IS" BASIS,
+ * WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+ * See the License for the specific language governing permissions and
+ * limitations under the License.
+ */
+
 package production;
 
 import org.hamcrest.CoreMatchers;
